---
title: "D19200 - tumornofibmono"
author: "Kai"
date: "13th January 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(ggplot2)

#create theme to not duplicate code
theme_jb <- function(base_size=11) {
  (theme_bw() +
     theme(
       axis.title = element_text(size=11),
       axis.text = element_text(size=11),
       legend.text = element_text(size=11),
       #optional rotation of x-axis labels
       axis.text.x = element_text(angle=90, hjust=1,vjust=.5)
     ))
}

theme_jb_nogrid <- function(base_size=11) {
  (theme_bw() +
     theme(
       axis.title = element_text(size=11),
       axis.text = element_text(size=11),
       legend.text = element_text(size=11),
       #optional rotation of x-axis labels
       axis.text.x = element_text(angle=90, hjust=1,vjust=.5),
       panel.grid = element_blank(),
       panel.border = element_blank(),
       axis.line = element_line()
     ))
}
```

# Load seurat objects

```{r}
OUTPUT = "~/tscr/kai/projects/01_SCC_scRNA/results/samples/cohort/D19200/rp/"
load(paste0(OUTPUT,"RData/D19200-tumor.RData"))
load(paste0(OUTPUT,"RData/D19200-filt.RData"))
```

# Subset Cells

```{r}
# remove cells that are probably fibroblasts and monocytes within the tumor cluster
tumorcells <- names(D19200.tumor$seurat_clusters_tumor_SCTcc_PCA_new[!as.character(D19200.tumor$seurat_clusters_tumor_SCTcc_PCA_new) %in% c("8","9")])
D19200.tumor <- subset(D19200.filt,cells = tumorcells)
```

# Normalization

After subsetting of cells we have to re-run the preprocessing steps for Normalization, dimensionality reduction, etc. 
This is needed, since the gene expression matrix is now reduced and enables more accurate capturing of the heterogeneity in the cells.
Important: The slots, reductions and SCT assay saved in D19200.tumor are still from the non-subsetted object, therefore we will override these in the next steps.

## Normalization of RNA assay

```{r SCC_normalization}
#RNA normalization without regressing, necessary for example for differential gene expression not based on SCT.
DefaultAssay(D19200.tumor) <- "RNA"
#log-normalization for slot "data"
D19200.tumor <- NormalizeData(D19200.tumor, 
                              normalization.method = "LogNormalize", #default
                              scale.factor = 10000, #default, scale factor 10000 is recommendation from seurat tutorial
                              assay = "RNA",
                              margin = 1 # default; normalizes across features
) 

#Find Variable Features
D19200.tumor <- FindVariableFeatures(D19200.tumor, 
                                     assay = "RNA",
                                     selection.method = "vst", #default
                                     nfeatures = 2000 #default; only 2000 , 3000 for SCT promises since to "restore" some more VariableFeatures
)

#scaling for slot "scale.data"
allgenes <- rownames(D19200.tumor)
D19200.tumor <- ScaleData(D19200.tumor, 
                          features = allgenes, 
                          do.scale = T,
                          do.center = T,
                          scale.max = 10,
                          assay = "RNA")
```

## Normalization with SCTransform

More information: https://doi.org/10.1186/s13059-019-1874-1

```{r SCC_normalization}
D19200.tumor <- SCTransform(D19200.tumor, 
                            assay="RNA", #default
                            new.assay.name = "SCT", #default; overwrites old SCT
                            do.correct.umi = T, #default change counts slot to corrected counts in new assay
                            variable.features.n = 3000, #default set variable features
                            vars.to.regress = NULL, #default optional variables to regress out
                            do.scale = F, #default scale pearson residuals in scale.data slot to have unit variance
                            do.center = T, #default center pearson residuals in scale.data to have mean zero, needed for PCA
                            return.only.var.genes = T, #default scale.data.matrices output assay only contains variable genes
                            seed.use = 1448145)
```

# Dimension reduction

## PCA

always based on SCT assay

```{r}
D19200.tumor <- RunPCA(D19200.tumor, 
                       npcs = 50,  #number of PCs to use
                       assay = "SCT",
                       rev.pca = F, # default, Run cell x gene matrix
                       weight.by.var = T, #default, Weight cell embedding by variance of each PC!)
                       approx = T, #if TRUE, uses irlba instead of prcomp, which is a single value decomposition to approximate PCs, saving computation time (by not calculating all PCs; important for big datasets) and being similar/same to PCs
                       features = VariableFeatures(D19200.tumor,assay = "SCT"), #default
                       reduction.name = "pca_SCT",
                       verbose = F)
```

```{r}
#calculate the % variance explained
PCAvar.SCT <- D19200.tumor@reductions$pca_SCT@stdev^2 / sum(matrixStats::rowVars(as.matrix(D19200.tumor@assays$SCT@scale.data)[VariableFeatures(D19200.tumor,assay = "SCT"),]))

PCAplot.SCT <- DimPlot(D19200.tumor, reduction = "pca_SCT", dims = c(1,2), group.by = "orig.ident") + 
  labs(x=paste0("principal component 1 (",signif(PCAvar.SCT[1]*100,digits = 2),"% variance)"),
       y=paste0("principal component 2 (",signif(PCAvar.SCT[2]*100,digits = 2),"% variance)")) +
  theme_jb()

PCAheatmap.SCT <- DimHeatmap(D19200.tumor, 
                             dims = 1:20, 
                             balanced = TRUE, 
                             fast=F, 
                             reduction = "pca_SCT",
                             assays = "SCT",
                             slot = "scale.data",combine = F) 
PCAheatmap.SCT <- lapply(PCAheatmap.SCT, function(x) x + theme_jb() + theme(axis.text.x = element_blank(), axis.line.x = element_blank()))

PCAelbowplot.SCT <- ggplot(data=data.frame(var=PCAvar.SCT,
                                           PC=1:length(PCAvar.SCT))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=var)) +
  labs(x="principal component", y="% variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

# cumulative Variance Plot
PCAsumplot.SCT <- ggplot(data=data.frame(cumvar=cumsum(PCAvar.SCT),PC=1:length(PCAvar.SCT))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=cumvar)) +
  labs(x="principal component", y="% cumulative variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()


ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-PCA.svg"),PCAplot.SCT, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-PCA-noleg.svg"),PCAplot.SCT + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
lapply(1:length(PCAheatmap.SCT), function(x) ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-PCA-heatmap-PC",x,".svg"),PCAheatmap.SCT[[x]], height=5, width=5))
lapply(1:length(PCAheatmap.SCT), function(x) ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-PCA-heatmap-PC",x,"-noleg.svg"),PCAheatmap.SCT[[x]] + theme(legend.position = "none"), height=5, width=5))
ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-PCA-elbow.svg"),PCAelbowplot.SCT, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-PCA-elbow-noleg.svg"),PCAelbowplot.SCT + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-PCA-variancesum.svg"),PCAsumplot.SCT, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-PCA-variancesum-noleg.svg"),PCAsumplot.SCT + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)

PCAplot.SCT
PCAheatmap.SCT
PCAelbowplot.SCT
PCAsumplot.SCT
```

## Graph-based clustering

here we perform two different clusterings, one based on the PCA of all cells, one of the tumor.

- seurat_clusters_all_SCT_PCA:      "old" seurat clusters before subsetting (from preprocessing script)
- seurat_clusters_tumor_SCT_PCA:    New seurat clusters from subsetting performed on PCA on SCT assay

```{r FindClusters}
numberofPCs = 20

D19200.tumor$seurat_clusters_all_SCT_PCA <- D19200.tumor$seurat_clusters
D19200.tumor$seurat_clusters <- NULL 
  
#based on PCA
D19200.tumor <- FindNeighbors(D19200.tumor, 
                                     dims = 1:numberofPCs, 
                                     reduction = "pca_SCT",
                                     assay = "SCT")
D19200.tumor <- FindClusters(D19200.tumor, 
                                    resolution = 0.8, #default 
                                    random.seed=100)

D19200.tumor$seurat_clusters_tumor_SCT_PCA <- D19200.tumor$seurat_clusters
D19200.tumor$seurat_clusters <- NULL 

#save new PCA with colored clusters 
PCAplot.SCT.clus <- DimPlot(D19200.tumor,
                            dims=1:2, 
                            reduction="pca_SCT", 
                            pt.size = 1, 
                            group.by = "seurat_clusters_tumor_SCT_PCA") + 
         labs(x=paste0("principal component 1 (",signif(PCAvar.SCT[1]*100,digits = 2),"% of variance)"),
              y=paste0("principal component 2 (",signif(PCAvar.SCT[2]*100,digits = 2),"% of variance)")) +
  theme_jb_nogrid()
ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-PCA-clus.svg"),PCAplot.SCT.clus, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-PCA-clus-noleg.svg"),PCAplot.SCT.clus + theme(legend.position = "none") + labs(title = NULL) + ggtitle(NULL), width = 5, height = 5)
PCAplot.SCT.clus
```

## UMAP

```{r UMAP, fig.width=10, fig.height=8}
D19200.tumor <- RunUMAP(D19200.tumor,
                        dims = 1:numberofPCs,
                        assay = "SCT",
                        umap.method = "uwot", # Seurat default
                        graph=NULL, #default
                        reduction="pca_SCT",
                        reduction.name = "umap_SCT"
)

UMAP.SCT.clusPCA <- DimPlot(D19200.tumor, 
                            reduction = "umap_SCT", 
                            pt.size = 1, 
                            label = T, 
                            group.by = "seurat_clusters_tumor_SCT_PCA") + theme(legend.position = "bottom") +
  theme_jb_nogrid()

ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-UMAP-clusPCA.svg"),UMAP.SCT.clusPCA)
ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-SCT-UMAP-clusPCA-noleg.svg"),UMAP.SCT.clusPCA + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
UMAP.SCT.clusPCA
```

## Save 

```{r}
#Single Seurat object
save(file = paste0(OUTPUT,"RData/D19200-tumornofibmono.RData"),D19200.tumor)
#SessionInfo
writeLines(capture.output(sessionInfo()), paste0(OUTPUT,"RData/03-D19200-tumornofibmono-sessionInfo.txt"))
```
