---
title: "D19200 - dorothea"
author: "Kai"
date: "13th January 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(ggplot2)

#create theme to not duplicate code
theme_jb <- function(base_size=11) {
  (theme_bw() +
     theme(
       axis.title = element_text(size=11),
       axis.text = element_text(size=11),
       legend.text = element_text(size=11),
       #optional rotation of x-axis labels
       axis.text.x = element_text(angle=90, hjust=1,vjust=.5),
       panel.grid.minor = element_blank()
     ))
}

theme_jb_nogrid <- function(base_size=11) {
  (theme_bw() +
     theme(
       axis.title = element_text(size=11),
       axis.text = element_text(size=11),
       legend.text = element_text(size=11),
       #optional rotation of x-axis labels
       axis.text.x = element_text(angle=90, hjust=1,vjust=.5),
       panel.grid = element_blank(),
       panel.border = element_blank(),
       axis.line = element_line()
     ))
}
```

# Load Seurat Object from preprocessing

```{r}
OUTPUT = "~/tscr/kai/projects/01_SCC_scRNA/results/samples/cohort/D19200/rp/"
load(paste0(OUTPUT,"RData/D19200-tumornofibmono.RData"))
```

# Change seurat clusters to new numbers

necessary for Violin Plots; numbers according to appaerance in trajectory

```{r}
D19200.tumor$trajectorycluster <- as.character(D19200.tumor$seurat_clusters_tumor_SCT_PCA)
D19200.tumor$trajectorycluster[D19200.tumor$trajectorycluster=="0"] <- "6 temp"
D19200.tumor$trajectorycluster[D19200.tumor$trajectorycluster=="1"] <- "3 temp"
D19200.tumor$trajectorycluster[D19200.tumor$trajectorycluster=="2"] <- "4 temp"
D19200.tumor$trajectorycluster[D19200.tumor$trajectorycluster=="3"] <- "2 temp"
D19200.tumor$trajectorycluster[D19200.tumor$trajectorycluster=="4"] <- "1 temp"
D19200.tumor$trajectorycluster[D19200.tumor$trajectorycluster=="5"] <- "7 temp"
D19200.tumor$trajectorycluster[D19200.tumor$trajectorycluster=="6"] <- "5 temp"
D19200.tumor$trajectorycluster <- sub(" temp","",D19200.tumor$trajectorycluster)
D19200.tumor$trajectorycluster <- factor(D19200.tumor$trajectorycluster)

clustercolors <- scales::hue_pal()(length(unique(D19200.tumor$trajectorycluster)))
names(clustercolors) <- c("6","3","4","2","1","7","5")

D19200.tumor <- CellCycleScoring(D19200.tumor,
                                 s.features = cc.genes$s.genes,
                                 g2m.features = cc.genes$g2m.genes,
                                 assay="RNA")
```

# DoRothEA

## Load regulons and expression data

VIPER with Dorothea on all tumor cells

```{r}
library(dorothea)
DefaultAssay(D19200.tumor) <- "SCT"
regulon <- get(data(dorothea_hs, package = "dorothea"))

set.seed(42)
D19200.tumor <- run_viper(D19200.tumor, regulon,
                          options = list(method = "scale",
                                         pleiotropy = F, # default
                                         nes = T, #default
                                         bootstraps = 0, #default
                                         minsize = 4, 
                                         eset.filter = F,
                                         cores = 4, 
                                         verbose = T),
                          assay_key = "SCT")

#Define TFs based on confidence level
DefaultAssay(D19200.tumor) <- "dorothea"

allTFs <- rownames(D19200.tumor@assays$dorothea)
regulonconf <- unique(as.data.frame(regulon)[,1:2])
regulonconf <- regulonconf[match(allTFs,regulonconf$tf),]

TFs <- list(ABCDE=rownames(D19200.tumor@assays$dorothea),
            ABCD=regulonconf[regulonconf$confidence %in% c("A","B","C","D"),]$tf,
            ABC=regulonconf[regulonconf$confidence %in% c("A","B","C"),]$tf,
            AB=regulonconf[regulonconf$confidence %in% c("A","B"),]$tf,
            A=regulonconf[regulonconf$confidence == "A",]$tf)

EMTTFs <- c("ZEB1","ZEB2","SNAI1","SNAI2","TWIST1","TWIST2","HIF1A")
```

Target overlap
```{r}
EMTTFtargets <- lapply(EMTTFs,function(x) regulon[regulon$tf %in% x,]$target)
names(EMTTFtargets) <- EMTTFs

targetoverlap <- lapply(EMTTFtargets,function(y) data.frame(t(unlist(lapply(EMTTFtargets,function(x) length(intersect(y,x)) / length(y))))))
targetoverlap <- as.data.frame(data.table::rbindlist(targetoverlap))
rownames(targetoverlap) <- names(targetoverlap)

#size of targets
library(ComplexHeatmap)
rightanno <- HeatmapAnnotation(size=unlist(lapply(EMTTFtargets,length)),
                               size_text=anno_text(unlist(lapply(EMTTFtargets,length))),
                               col = list(size=circlize::colorRamp2(breaks=c(0,max(unlist(lapply(EMTTFtargets,length)))),colors = c("white","blue"))),
                               which="row", annotation_name_side = "bottom",annotation_name_gp = gpar(fontsize=11))
bottomanno <- HeatmapAnnotation(size=unlist(lapply(EMTTFtargets,length)),
                                size_text=anno_text(unlist(lapply(EMTTFtargets,length))),
                                col = list(size=circlize::colorRamp2(breaks=c(0,max(unlist(lapply(EMTTFtargets,length)))),colors = c("white","blue"))),
                                which="column", annotation_name_side = "left",annotation_name_gp = gpar(fontsize=11))

#relative to rows (i.e., EMT TF in row is 100% of targets)
tov <- Heatmap(targetoverlap*100,
               name="percentage of overlapping targets",
               column_title = "EMT TF",
               row_title = "EMT TF (relative to rows, i.e., rows are 100% or targets)",
               column_title_side = "bottom",
               col=circlize::colorRamp2(breaks=c(0,100), colors = c("white","red")),
               cluster_columns = F,
               cluster_rows = F,
               cell_fun = function(j, i, x, y, width, height, fill) {
                 grid.text(sprintf("%.0f", (targetoverlap*100)[i, j]), x, y, gp = gpar(fontsize = 10))},
               right_annotation = rightanno,
               bottom_annotation = bottomanno,
               #graphic options
               row_names_gp = gpar(fontsize=11),
               column_names_gp = gpar(fontsize=11),
               row_title_gp = gpar(fontsize=11),
               heatmap_legend_param = list(title_gp = gpar(fontsize=11), labels_gp = gpar(fontsize=11)),
               border = T,
               use_raster = T,
               raster_device = "CairoPNG",
               raster_quality = 4,
               width = unit(5,"in"),
               height = unit(5,"in"))

pdf(paste0(OUTPUT,"figs-D19200/D19200-EMTTF-targetsfromdorothea-ov.pdf"), width=20, height=20)
tov
dev.off()
```

## Summarize cluster

```{r}
library(tidyr)
library(tibble)
library(dplyr)
## Summarise cluster
## We transform Viper scores, scaled by seurat, into a data frame to better 
## handling the results
viper_scores_df <- GetAssayData(D19200.tumor, 
                                slot = "data", 
                                assay = "dorothea") %>%
  data.frame(check.names = F) %>%
  t()

## We create a data frame containing the cells and their clusters
CellsClusters <- data.frame(cell = names(D19200.tumor$trajectorycluster), 
                            cell_type = as.character(D19200.tumor$trajectorycluster),
                            check.names = F)

## We create a data frame with the Viper score per cell and its clusters
viper_scores_clusters <- viper_scores_df  %>%
  data.frame() %>% 
  rownames_to_column("cell") %>%
  gather(tf, activity, -cell) %>%
  inner_join(CellsClusters)

## We summarize the Viper scores by cellpopulation
summarized_viper_scores <- viper_scores_clusters %>% 
  group_by(tf, cell_type) %>%
  summarise(avg = mean(activity),
            std = sd(activity))

dtsum.df <- reshape2::dcast(as.data.frame(summarized_viper_scores),tf ~ cell_type,value.var = "avg")
rownames(dtsum.df) <- dtsum.df[,1]
dtsum.df <- dtsum.df[,-1]
```

## Investigate EMT transcription factors

### Visualization as Heatmap with each cell sorted by pseudotime

```{r}
library(RColorBrewer)
library(ComplexHeatmap)
D19200.tumor.temp <- D19200.tumor
load(paste0(OUTPUT,"RData/D19200-tumornofibmono-pseudotimeseurat.RData"))
ptcrv1 <- D19200.tumor$pca_SCT.crv1
ptcrv2 <- D19200.tumor$pca_SCT.crv2
D19200.tumor <- D19200.tumor.temp

dtEMT <- D19200.tumor@assays$dorothea@data[EMTTFs,]
colorbreaks.dtEMT <- c(min(dtEMT,na.rm = T),median(dtEMT[dtEMT<0]),0,median(dtEMT[dtEMT>0]),max(dtEMT,na.rm = T))

#crv1
#same ordering as heatmap data
ptcrv1 <- ptcrv1[colnames(dtEMT)]
bottomanno.ptcrv1 <- HeatmapAnnotation(df = data.frame(cluster=D19200.tumor$trajectorycluster[!is.na(ptcrv1)],
                                                       Phase=D19200.tumor$Phase[!is.na(ptcrv1)],
                                                       nCount=D19200.tumor$nCount_RNA_log10[!is.na(ptcrv1)],
                                                       nFeature=D19200.tumor$nFeature_RNA_log10[!is.na(ptcrv1)],
                                                       mito=D19200.tumor$percent.mito[!is.na(ptcrv1)],
                                                       hkgenes=D19200.tumor$n.exp.hkgenes[!is.na(ptcrv1)]),
                                       col = list(cluster = clustercolors,
                                                  nCount = circlize::colorRamp2(breaks=c(0,max(D19200.tumor$nCount_RNA_log10)),colors = c("black","yellow")),
                                                  nFeature = circlize::colorRamp2(breaks=c(0,max(D19200.tumor$nFeature_RNA_log10)),colors = c("black","yellow")),
                                                  mito = circlize::colorRamp2(breaks=c(0,max(D19200.tumor$percent.mito)),colors = c("black","yellow")),
                                                  hkgenes = circlize::colorRamp2(breaks=c(0,max(D19200.tumor$n.exp.hkgenes)),colors = c("black","yellow")),
                                                  Phase = c(G1=brewer.pal(3,"Set1")[1],
                                                            G2M=brewer.pal(3,"Set1")[2],
                                                            S=brewer.pal(3,"Set1")[3])),
                                       which = "column", annotation_name_side = "left")

dtEMT.ptcrv1.hm <- Heatmap(dtEMT[,names(ptcrv1)[!is.na(ptcrv1)]],
                           name="dorothea data",
                           col=circlize::colorRamp2(breaks=colorbreaks.dtEMT, colors = c("blue","purple", "black", "yellow","orange")),
                           show_column_names = F,
                           show_row_names = T,
                           cluster_columns = F,
                           #column_split = is.na(ptcrv1),
                           column_order = names(ptcrv1[!is.na(ptcrv1)][order(ptcrv1[!is.na(ptcrv1)])]),
                           cluster_rows = F,
                           bottom_annotation = bottomanno.ptcrv1, 
                           column_title_rot = 90,
                           #column_gap = unit(1,"cm"),
                           #graphic options
                           row_names_gp = gpar(fontsize=11),
                           column_names_gp = gpar(fontsize=11),
                           row_title_gp = gpar(fontsize=11),
                           heatmap_legend_param = list(title_gp = gpar(fontsize=11), labels_gp = gpar(fontsize=11)),
                           border = T,
                           width = unit(5,"in"),
                           height = unit(2.5,"in"),
                           use_raster = T,
                           raster_device = "CairoPNG",
                           raster_quality = 4
)

#crv 2
#same ordering as heatmap data
ptcrv2 <- ptcrv2[colnames(dtEMT)]

#new bottom annotation
bottomanno.ptcrv2 <- HeatmapAnnotation(df = data.frame(cluster=D19200.tumor$trajectorycluster[!is.na(ptcrv2)],
                                                       Phase=D19200.tumor$Phase[!is.na(ptcrv2)],
                                                       nCount=D19200.tumor$nCount_RNA_log10[!is.na(ptcrv2)],
                                                       nFeature=D19200.tumor$nFeature_RNA_log10[!is.na(ptcrv2)],
                                                       mito=D19200.tumor$percent.mito[!is.na(ptcrv2)],
                                                       hkgenes=D19200.tumor$n.exp.hkgenes[!is.na(ptcrv2)]),
                                       col = list(cluster = clustercolors,
                                                  nCount = circlize::colorRamp2(breaks=c(0,max(D19200.tumor$nCount_RNA_log10)),colors = c("black","yellow")),
                                                  nFeature = circlize::colorRamp2(breaks=c(0,max(D19200.tumor$nFeature_RNA_log10)),colors = c("black","yellow")),
                                                  mito = circlize::colorRamp2(breaks=c(0,max(D19200.tumor$percent.mito)),colors = c("black","yellow")),
                                                  hkgenes = circlize::colorRamp2(breaks=c(0,max(D19200.tumor$n.exp.hkgenes)),colors = c("black","yellow")),
                                                  Phase = c(G1=brewer.pal(3,"Set1")[1],
                                                            G2M=brewer.pal(3,"Set1")[2],
                                                            S=brewer.pal(3,"Set1")[3])),
                                       which = "column", annotation_name_side = "left")


dtEMT.ptcrv2.hm <- Heatmap(dtEMT[,names(ptcrv2)[!is.na(ptcrv2)]],
                           name="dorothea data",
                           col=circlize::colorRamp2(breaks=colorbreaks.dtEMT, colors = c("blue","purple", "black", "yellow","orange")),
                           show_column_names = F,
                           show_row_names = T,
                           cluster_columns = F,
                           #column_split = is.na(ptcrv1),
                           column_order = names(ptcrv2[!is.na(ptcrv2)][order(ptcrv2[!is.na(ptcrv2)])]),
                           cluster_rows = F,
                           bottom_annotation = bottomanno.ptcrv2, 
                           column_title_rot = 90,
                           #column_gap = unit(1,"cm"),
                           #graphic options
                           row_names_gp = gpar(fontsize=11),
                           column_names_gp = gpar(fontsize=11),
                           row_title_gp = gpar(fontsize=11),
                           heatmap_legend_param = list(title_gp = gpar(fontsize=11), labels_gp = gpar(fontsize=11)),
                           border = T,
                           width = unit(5,"in"),
                           height = unit(2.5,"in"),
                           use_raster = T,
                           raster_device = "CairoPNG",
                           raster_quality = 4
)

pdf(paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-dtEMT-HM-clusPCA-crv1.pdf"), width=10, height=10)
dtEMT.ptcrv1.hm
dev.off()

pdf(paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-dtEMT-HM-clusPCA-crv2.pdf"), width=10, height=10)
dtEMT.ptcrv2.hm
dev.off()
```

### Visualization as Heatmap with summarised activity per cluster

```{r}
EMTdtsum.df <- dtsum.df[EMTTFs,]

EMTdtsum.df.cb <- c(-max(c(abs(min(EMTdtsum.df)),abs(max(EMTdtsum.df)))),
                    -max(c(abs(min(EMTdtsum.df)),abs(max(EMTdtsum.df))))/2,
                    0,
                    max(c(abs(min(EMTdtsum.df)),abs(max(EMTdtsum.df))))/2,
                    max(c(abs(min(EMTdtsum.df)),abs(max(EMTdtsum.df)))))

EMTdtsum.df.cb <- c(-max(c(abs(min(EMTdtsum.df)),abs(max(EMTdtsum.df)))),0,max(c(abs(min(EMTdtsum.df)),abs(max(EMTdtsum.df)))))

EMTdtsum.df.hm.crv1 <- Heatmap(EMTdtsum.df[,1:5],
                               name="dorothea sum",
                               col=circlize::colorRamp2(breaks=EMTdtsum.df.cb, colors = c("purple", "black", "yellow")),
                               show_column_names = T,
                               show_row_names = T,
                               cluster_columns = F,
                               cluster_rows = F,
                               #graphic options
                               row_names_gp = gpar(fontsize=11),
                               column_names_gp = gpar(fontsize=11),
                               row_title_gp = gpar(fontsize=11),
                               heatmap_legend_param = list(title_gp = gpar(fontsize=11), labels_gp = gpar(fontsize=11)),
                               border = T,
                               heatmap_width = unit(4,"in"),
                               heatmap_height = unit(4,"in"),
                               use_raster = T,
                               raster_device = "CairoPNG",
                               raster_quality = 4
)

EMTdtsum.df.hm.crv2 <- Heatmap(EMTdtsum.df[,c(1:3,6,7)],
                               name="dorothea sum",
                               col=circlize::colorRamp2(breaks=EMTdtsum.df.cb, colors = c("purple", "black", "yellow")),
                               show_column_names = T,
                               show_row_names = T,
                               cluster_columns = F,
                               cluster_rows = F,
                               #graphic options
                               row_names_gp = gpar(fontsize=11),
                               column_names_gp = gpar(fontsize=11),
                               row_title_gp = gpar(fontsize=11),
                               heatmap_legend_param = list(title_gp = gpar(fontsize=11), labels_gp = gpar(fontsize=11)),
                               border = T,
                               heatmap_width = unit(4,"in"),
                               heatmap_height = unit(4,"in"),
                               use_raster = T,
                               raster_device = "CairoPNG",
                               raster_quality = 4
)

pdf(paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-dtEMTsum-HM-clusPCA-crv1.pdf"), width=10, height=10)
EMTdtsum.df.hm.crv1
dev.off()
pdf(paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-dtEMTsum-HM-clusPCA-crv2.pdf"), width=10, height=10)
EMTdtsum.df.hm.crv2
dev.off()

```

### Visualizations as Violinplot

```{r}
dtEMT.m <- reshape2::melt(dtEMT)
names(dtEMT.m) <- c("TF","cell","activity")
dtEMT.m$Phase <- D19200.tumor$Phase[match(dtEMT.m$cell,names(D19200.tumor$Phase))]
dtEMT.m$trajectorycluster <- D19200.tumor$trajectorycluster[match(dtEMT.m$cell,names(D19200.tumor$trajectorycluster))]

#Visualize as Violin plot
dtEMT.vln <- ggplot(data=dtEMT.m, aes(x=TF,y=activity,fill=trajectorycluster)) +
  geom_boxplot(position="dodge",outlier.size = 0.5) +
  scale_y_continuous(breaks = seq(-10,10,1)) +
  scale_fill_manual(values = clustercolors) +
  geom_hline(yintercept = 0,col="black") +
  labs(x="", y="TF activity") +
  theme_jb()

ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-dtEMT-VLN-clusPCA.svg"),dtEMT.vln,width = 6,height = 5)
ggsave(filename = paste0(OUTPUT,"figs-D19200/D19200-tumornofibmono-dtEMT-VLN-clusPCA-noleg.svg"),dtEMT.vln + theme(legend.position = "none") + ggtitle(""),width = 8,height = 5.5)
```

# Save 

```{r}
#save seurat object
save(file = paste0(OUTPUT,"RData/D19200-tumornofibmono-dorothea.RData"),D19200.tumor)
#save SessionInfo
writeLines(capture.output(sessionInfo()), paste0(OUTPUT,"RData/08-D19200-tumornofibmono-dorothea-sessionInfo.txt"))
```
