---
title: "Merge cells"
author: "Kai"
date: "6th April 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(ggplot2)

source(paste0("~/tscr/kai/projects/01_SCC_scRNA/results/functions.R"))
OUTPUT = "~/tscr/kai/projects/01_SCC_scRNA/results/alldatasets/"
```

# load all seurat objects

```{r}
#tumor cells
#LN-MET
load("~/tscr/kai/projects/01_SCC_scRNA/results/samples/cohort/patient1/RData/patient1-tumorfilt-typeanalysis.RData")
#cohort
load("~/tscr/kai/projects/01_SCC_scRNA/results/samples/cohort/all/rp/RData/SCCfilt-epino1filt.RData")
#Kürten
load("~/tscr/kai/projects/01_SCC_scRNA/results/pubdatasets/kuertendata/HPVneg_CD45n_analysis/RData/HNSCC-epifilt.RData")
```

# Merge all Cells

```{r}
patient1.tumor.new = CreateSeuratObject(counts = patient1.tumor@assays$RNA@counts)
patient1.tumor.new$sample <- "1"
patient1.tumor.new$patients <- "1"
patient1.tumor.new$tumortype <- patient1.tumor$tumortype
patient1.tumor.new$tumortype_ov <- patient1.tumor$tumortype_ov
patient1.tumor.new$Lok <- "LN-MET"
HNSCC.e.new = CreateSeuratObject(counts = HNSCC.e@assays$RNA@counts)
HNSCC.e.new$sample <- HNSCC.e$patients
HNSCC.e.new$patients <- HNSCC.e$patients
HNSCC.e.new$Lok <- "PT"
SCC.epino1.new = CreateSeuratObject(counts = SCC.epino1@assays$RNA@counts)
SCC.epino1.new$sample <- SCC.epino1$sample
SCC.epino1.new$patients <- SCC.epino1$patient
SCC.epino1.new$Lok <- SCC.epino1$Lok

SCC <- merge(x = patient1.tumor.new,
             y = list(HNSCC.e.new,
                      SCC.epino1.new))

SCC$patients <- factor(SCC$patients,levels = sort(unique(SCC$patients)))
SCC$Lok[SCC$Lok=="LN-MET"] <- "MET"
SCC$Lok <- factor(SCC$Lok,levels=c("PT","MET","NA"))

rm(patient1.tumor,patient1.tumor.new)
rm(SCC.epino1,SCC.epino1.new)
rm(HNSCC.e,HNSCC.e.new)
```

# Normalization

## Normalization of RNA assay

```{r SCC_normalization}
#RNA normalization without regressing, necessary for example for differential gene expression not based on SCT.
DefaultAssay(SCC) <- "RNA"
#log-normalization for slot "data"
SCC <- NormalizeData(SCC, 
                     normalization.method = "LogNormalize", #default
                     scale.factor = 10000, #default, scale factor 10000 is recommendation from seurat tutorial
                     assay = "RNA",
                     margin = 1 # default; normalizes across features
) 

#Find Variable Features
SCC <- FindVariableFeatures(SCC, 
                            assay = "RNA",
                            selection.method = "vst", #default
                            nfeatures = 2000 #default; only 2000 , 3000 for SCT promises since to "restore" some more VariableFeatures
)

#scaling for slot "scale.data"
SCC <- ScaleData(SCC,
                 features = VariableFeatures(SCC,assay = "RNA"),
                 do.scale = T,
                 do.center = T,
                 scale.max = 10,
                 assay = "RNA")
```

## Determine cell cycle

```{r}
SCC$nCount_RNA_log10 <- log10(SCC$nCount_RNA)
SCC$nFeature_RNA_log10 <- log10(SCC$nFeature_RNA)

#from Tirosh et al 2015 cell cycle genes, included in Seurat package
#The "data" slot is always picked, based on RNA assay
SCC <- CellCycleScoring(SCC,
                        s.features = cc.genes.updated.2019$s.genes,
                        g2m.features = cc.genes.updated.2019$g2m.genes, 
                        assay="RNA")

#Assign the expression of mitochondrial and housekeeping genes
#houskeeping genes are e.g. ribosomal proteins, expressed in nearly every cell and less sensitive to high dropout.
# relative expression of mitochondrial gene as metric for damaged cells
SCC[["percent.mito"]] <- PercentageFeatureSet(SCC, pattern = "^MT-") 
#read in names of housekeeper genes
hkgenes <- read.table("~/tscr/kai/projects/01_SCC_scRNA/db/tirosh_house_keeping.txt", skip = 2)
hkgenes <- as.vector(hkgenes$V1) 
# remove hkgenes that were not found (maybe not in the reference or got lost)
hkgenes.found <- which(toupper(rownames(SCC)) %in% toupper(hkgenes))
# sum of hk genes for each cell
hk_per_cell <- Matrix::colSums(GetAssayData(SCC)[hkgenes.found, ] > 0)
SCC <- AddMetaData(object = SCC, metadata = hk_per_cell, col.name = "n.exp.hkgenes")
rm(hk_per_cell,hkgenes.found)
```

## Normalization with SCTransform

More information: https://doi.org/10.1186/s13059-019-1874-1

```{r SCC_normalization}
SCC <- SCTransform(SCC, 
                   ncells = ncol(SCC),
                   assay="RNA", #default
                   new.assay.name = "SCT", #default; overwrites old SCT
                   do.correct.umi = T, #default change counts slot to corrected counts in new assay
                   variable.features.n = 3000, #default set variable features
                   vars.to.regress = NULL, #default optional variables to regress out
                   do.scale = F, #default scale pearson residuals in scale.data slot to have unit variance
                   do.center = T, #default center pearson residuals in scale.data to have mean zero, needed for PCA
                   return.only.var.genes = T, #default scale.data.matrices output assay only contains variable genes
                   seed.use = 1448145)
```

## Normalization with SCTransform and regression

More information: https://doi.org/10.1186/s13059-019-1874-1

```{r SCC_normalization}
SCC <- SCTransform(SCC, 
                   ncells = ncol(SCC),
                   assay="RNA", #default
                   new.assay.name = "SCTccmito", #default; overwrites old SCT
                   do.correct.umi = T, #default change counts slot to corrected counts in new assay
                   variable.features.n = 3000, #default set variable features
                   vars.to.regress = c("S.Score","G2M.Score","percent.mito"), #default optional variables to regress out
                   do.scale = F, #default scale pearson residuals in scale.data slot to have unit variance
                   do.center = T, #default center pearson residuals in scale.data to have mean zero, needed for PCA
                   return.only.var.genes = T, #default scale.data.matrices output assay only contains variable genes
                   seed.use = 1448145)
```

# Dimension reduction

## PCA

always based on SCT assay

```{r}
SCC <- RunPCA(SCC, 
              npcs = 50,  #number of PCs to use
              assay = "SCT",
              rev.pca = F, # default, Run cell x gene matrix
              weight.by.var = T, #default, Weight cell embedding by variance of each PC!)
              approx = T, #if TRUE, uses irlba instead of prcomp, which is a single value decomposition to approximate PCs, saving computation time (by not calculating all PCs; important for big datasets) and being similar/same to PCs
              features = VariableFeatures(SCC,assay = "SCT"), #default
              reduction.name = "pca_SCT",
              verbose = F)
```

```{r}
#calculate the % variance explained
PCAvar.SCT <- SCC@reductions$pca_SCT@stdev^2 / sum(matrixStats::rowVars(as.matrix(SCC@assays$SCT@scale.data)[VariableFeatures(SCC,assay = "SCT"),]))

PCAplot.SCT <- DimPlot(SCC, reduction = "pca_SCT", dims = c(1,2), group.by = "patients") + 
  labs(x=paste0("principal component 1 (",signif(PCAvar.SCT[1]*100,digits = 2),"% of variance)"),
       y=paste0("principal component 2 (",signif(PCAvar.SCT[2]*100,digits = 2),"% of variance)")) +
  theme_jb()

PCAelbowplot.SCT <- ggplot(data=data.frame(var=PCAvar.SCT,
                                           PC=1:length(PCAvar.SCT))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=var)) +
  labs(x="principal component", y="% of variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

#cumulative variance plot
PCAsumplot.SCT <- ggplot(data=data.frame(cumvar=cumsum(PCAvar.SCT),PC=1:length(PCAvar.SCT))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=cumvar)) +
  labs(x="principal component", y="% of cumulative variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-PCA.svg"),PCAplot.SCT, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-PCA-noleg.svg"),PCAplot.SCT + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-PCA-elbow.svg"),PCAelbowplot.SCT, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-PCA-elbow-noleg.svg"),PCAelbowplot.SCT + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-PCA-variancesum.svg"),PCAsumplot.SCT, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-PCA-variancesum-noleg.svg"),PCAsumplot.SCT + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)

PCAplot.SCT
PCAelbowplot.SCT
PCAsumplot.SCT
```

## Graph-based clustering

```{r FindClusters}
numberofPCs = 20

DefaultAssay(SCC) <- "SCT"
#based on PCA
SCC <- FindNeighbors(SCC, 
                     dims = 1:numberofPCs, 
                     reduction = "pca_SCT",
                     assay = "SCT")
SCC <- FindClusters(SCC, 
                    resolution = 0.8,
                    random.seed=100)

SCC$seurat_clusters_all_SCT_PCA <- SCC$seurat_clusters
SCC$seurat_clusters <- NULL 
```

## UMAP

```{r UMAP, fig.width=10, fig.height=8}
SCC <- RunUMAP(SCC,
               dims = 1:numberofPCs,
               assay = "SCT",
               umap.method = "uwot", # Seurat default
               graph=NULL, #default
               reduction="pca_SCT",
               reduction.name = "umap_SCT"
)
```

```{r}
UMAP.SCT.clusPCA <- DimPlot(SCC, 
                            reduction = "umap_SCT", 
                            pt.size = 1, 
                            label = T, 
                            group.by = "seurat_clusters_all_SCT_PCA") +
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAP.SCT.patients <- DimPlot(SCC, 
                            reduction = "umap_SCT", 
                            pt.size = 1, 
                            label = F, 
                            group.by = "patients") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAP.SCT.Lok <- DimPlot(SCC, 
                        reduction = "umap_SCT", 
                        pt.size = 1, 
                        label = F, 
                        group.by = "Lok") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAP.SCT.QC <- FeaturePlot(SCC, 
                           features = c("nCount_RNA_log10", "nFeature_RNA_log10", "percent.mito", "n.exp.hkgenes","S.Score","G2M.Score"), 
                           reduction = "umap_SCT", 
                           cols = c("grey95", "blue"),
                           pt.size = 1,
                           order = T,
                           slot = "data")

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-UMAP-clusPCA.svg"),UMAP.SCT.clusPCA,width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-UMAP-patients.svg"),UMAP.SCT.patients,width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-UMAP-Lok.svg"),UMAP.SCT.Lok,width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-UMAP-QC.svg"),UMAP.SCT.QC,width = 5, height=5)

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-UMAP-clusPCA-noleg.svg"),UMAP.SCT.clusPCA + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-UMAP-patients-noleg.svg"),UMAP.SCT.patients + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-UMAP-Lok-noleg.svg"),UMAP.SCT.Lok + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCT-UMAP-QC-noleg.svg"),UMAP.SCT.QC + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)

UMAP.SCT.clusPCA
UMAP.SCT.patients
UMAP.SCT.Lok
UMAP.SCT.QC
```

# Dimension reduction with regression

## PCA

always based on SCTccmito assay

```{r}
SCC <- RunPCA(SCC, 
              npcs = 50,  #number of PCs to use
              assay = "SCTccmito",
              rev.pca = F, # default, Run cell x gene matrix
              weight.by.var = T, #default, Weight cell embedding by variance of each PC!)
              approx = T, #if TRUE, uses irlba instead of prcomp, which is a single value decomposition to approximate PCs, saving computation time (by not calculating all PCs; important for big datasets) and being similar/same to PCs
              features = VariableFeatures(SCC,assay = "SCTccmito"), #default
              reduction.name = "pca_SCTccmito",
              verbose = F)
```

```{r}
#calculate the % variance explained
PCAvar.SCTccmito <- SCC@reductions$pca_SCTccmito@stdev^2 / sum(matrixStats::rowVars(as.matrix(SCC@assays$SCTccmito@scale.data)[VariableFeatures(SCC,assay = "SCTccmito"),]))

PCAplot.SCTccmito <- DimPlot(SCC, reduction = "pca_SCTccmito", dims = c(1,2), group.by = "patients") + 
  labs(x=paste0("principal component 1 (",signif(PCAvar.SCTccmito[1]*100,digits = 2),"% of variance)"),
       y=paste0("principal component 2 (",signif(PCAvar.SCTccmito[2]*100,digits = 2),"% of variance)")) +
  theme_jb()

PCAelbowplot.SCTccmito <- ggplot(data=data.frame(var=PCAvar.SCTccmito,
                                           PC=1:length(PCAvar.SCTccmito))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=var)) +
  labs(x="principal component", y="% of variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

#cumulative variance plot
PCAsumplot.SCTccmito <- ggplot(data=data.frame(cumvar=cumsum(PCAvar.SCTccmito),PC=1:length(PCAvar.SCTccmito))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=cumvar)) +
  labs(x="principal component", y="% of cumulative variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-PCA.svg"),PCAplot.SCTccmito, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-PCA-noleg.svg"),PCAplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-PCA-elbow.svg"),PCAelbowplot.SCTccmito, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-PCA-elbow-noleg.svg"),PCAelbowplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-PCA-variancesum.svg"),PCAsumplot.SCTccmito, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-PCA-variancesum-noleg.svg"),PCAsumplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)

PCAplot.SCTccmito
PCAelbowplot.SCTccmito
PCAsumplot.SCTccmito
```

## Graph-based clustering

```{r FindClusters}
numberofPCs = 20

DefaultAssay(SCC) <- "SCTccmito"
#based on PCA
SCC <- FindNeighbors(SCC, 
                     dims = 1:numberofPCs, 
                     reduction = "pca_SCTccmito",
                     assay = "SCTccmito")
SCC <- FindClusters(SCC, 
                    resolution = 0.8,
                    random.seed=100)

SCC$seurat_clusters_all_SCTccmito_PCA <- SCC$seurat_clusters
SCC$seurat_clusters <- NULL 
```

## UMAP

```{r UMAP, fig.width=10, fig.height=8}
SCC <- RunUMAP(SCC,
               dims = 1:numberofPCs,
               assay = "SCTccmito",
               umap.method = "uwot", # Seurat default
               graph=NULL, #default
               reduction="pca_SCTccmito",
               reduction.name = "umap_SCTccmito"
)
```

```{r}
UMAP.SCTccmito.clusPCA <- DimPlot(SCC, 
                                  reduction = "umap_SCTccmito", 
                                  pt.size = 1, 
                                  label = T, 
                                  group.by = "seurat_clusters_all_SCTccmito_PCA") +
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAP.SCTccmito.patients <- DimPlot(SCC, 
                                   reduction = "umap_SCTccmito", 
                                   pt.size = 1, 
                                   label = F, 
                                   group.by = "patients") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAP.SCTccmito.Lok <- DimPlot(SCC, 
                              reduction = "umap_SCTccmito", 
                              pt.size = 1, 
                              label = F, 
                              group.by = "Lok") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAP.SCTccmito.QC <- FeaturePlot(SCC, 
                           features = c("nCount_RNA_log10", "nFeature_RNA_log10", "percent.mito", "n.exp.hkgenes","S.Score","G2M.Score"), 
                           reduction = "umap_SCTccmito", 
                           cols = c("grey95", "blue"),
                           pt.size = 1,
                           order = T,
                           slot = "data")

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAP-clusPCA.svg"),UMAP.SCTccmito.clusPCA,width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAP-patients.svg"),UMAP.SCTccmito.patients,width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAP-Lok.svg"),UMAP.SCTccmito.Lok,width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAP-QC.svg"),UMAP.SCTccmito.QC,width = 5, height=5)

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAP-clusPCA-noleg.svg"),UMAP.SCTccmito.clusPCA + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAP-patients-noleg.svg"),UMAP.SCTccmito.patients + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAP-Lok-noleg.svg"),UMAP.SCTccmito.Lok + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAP-QC-noleg.svg"),UMAP.SCTccmito.QC + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)

UMAP.SCTccmito.clusPCA
UMAP.SCTccmito.patients
UMAP.SCTccmito.Lok
UMAP.SCTccmito.QC
```

# Harmony batch correction

```{r}
library(harmony)
set.seed(42)
SCC <- RunHarmony(SCC,
                  reduction="pca_SCTccmito",
                  group.by.vars = "patients",
                  reduction.save = "harmony_SCTccmito",
                  plot_convergence = T,
                  max.iter.harmony = 30,
                  max.iter.cluster = 20,
                  assay.use = "SCTccmito",
                  project.dim = F)

set.seed(42)
SCC <- RunHarmony(SCC,
                  reduction="harmony_SCTccmito",
                  group.by.vars = "patients",
                  reduction.save = "harmony2_SCTccmito",
                  plot_convergence = T,
                  max.iter.harmony = 30,
                  max.iter.cluster = 20, 
                  assay.use = "SCTccmito",
                  project.dim = F)
```

## Graph-based clustering

```{r FindClusters}
numberofPCs = 20

DefaultAssay(SCC) <- "SCTccmito"
#based on PCA
SCC <- FindNeighbors(SCC, 
                     dims = 1:numberofPCs, 
                     reduction = "harmony2_SCTccmito",
                     assay = "SCTccmito")
SCC <- FindClusters(SCC, 
                    resolution = 0.8,
                    random.seed=100)

SCC$seurat_clusters_all_SCTccmito_harmony <- SCC$seurat_clusters
SCC$seurat_clusters <- NULL 
```

## UMAP

```{r UMAP, fig.width=10, fig.height=8}
SCC <- RunUMAP(SCC,
               dims = 1:numberofPCs,
               assay = "SCTccmito",
               umap.method = "uwot", # Seurat default
               graph=NULL, #default
               reduction="harmony2_SCTccmito",
               reduction.name = "humap_SCTccmito"
)
```

```{r}
UMAPharmony.SCTccmito.clusPCA <- DimPlot(SCC, 
                                         reduction = "humap_SCTccmito", 
                                         pt.size = 1, 
                                         label = T, 
                                         group.by = "seurat_clusters_all_SCTccmito_PCA") +
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAPharmony.SCTccmito.patients <- DimPlot(SCC, 
                                          reduction = "humap_SCTccmito", 
                                          pt.size = 1, 
                                          label = F, 
                                          group.by = "patients") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAPharmony.SCTccmito.Lok <- DimPlot(SCC, 
                                     reduction = "humap_SCTccmito", 
                                     pt.size = 1, 
                                     label = F, 
                                     group.by = "Lok") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAPharmony.SCTccmito.QC <- FeaturePlot(SCC, 
                                        features = c("nCount_RNA_log10", "nFeature_RNA_log10", "percent.mito", "n.exp.hkgenes","S.Score","G2M.Score"), 
                                        reduction = "humap_SCTccmito", 
                                        cols = c("grey95", "blue"),
                                        pt.size = 1,
                                        order = T,
                                        slot = "data")

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAPharmony-patients.svg"),UMAPharmony.SCTccmito.patients,width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAPharmony-Lok.svg"),UMAPharmony.SCTccmito.Lok,width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAPharmony-QC.svg"),UMAPharmony.SCTccmito.QC,width = 5, height=5)

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAPharmony-clusPCA-noleg.svg"),UMAPharmony.SCTccmito.clusPCA + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAPharmony-patients-noleg.svg"),UMAPharmony.SCTccmito.patients + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAPharmony-Lok-noleg.svg"),UMAPharmony.SCTccmito.Lok + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAPharmony-QC-noleg.svg"),UMAPharmony.SCTccmito.QC + theme(legend.position = "none") + ggtitle(NULL),width = 5, height=5)

UMAPharmony.SCTccmito.clusPCA
UMAPharmony.SCTccmito.patients
UMAPharmony.SCTccmito.Lok
UMAPharmony.SCTccmito.QC
```

## Graph-based clustering at different resolutions for harmony corrected

```{r FindClusters}
DefaultAssay(SCC) <- "SCTccmito"
numberofPCs = 20
#based on harmony
for (res in c(seq(1,4,1))) {
  SCC <- FindNeighbors(SCC,
                       dims = 1:numberofPCs,
                       reduction = "harmony2_SCTccmito",
                       assay = "SCTccmito")
  SCC <- FindClusters(SCC,
                      resolution = res,
                      random.seed=100)
  
  SCC[[paste0("seurat_clusters_all_SCTccmito_PCA_harmony_res",res)]] <- SCC$seurat_clusters
  SCC$seurat_clusters <- NULL
}

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAPharmony-clusPCAharmony-overview.svg"),
       ggplot() + geom_bar(stat="identity", 
                           mapping = aes(y=unlist(lapply(seq(1,4,1),function(res) length(table(SCC[[paste0("seurat_clusters_all_SCTccmito_PCA_harmony_res",res)]])))),x=seq(1,4,1))) +
         labs(x="resolution",y="number of clusters") +
         scale_y_continuous(breaks=seq(0,100,1)) +
         scale_x_continuous(breaks=seq(0,5,1)) +
         theme_jb(),
       width=5, height=5)

ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAPharmony-clusPCAharmony-overviewcells.svg"),
       ggplot() + geom_bar(stat="identity", 
                           mapping = aes(y=unlist(lapply(seq(1,4,1),function(res) mean(table(SCC[[paste0("seurat_clusters_all_SCTccmito_PCA_harmony_res",res)]])))),x=seq(1,4,1))) +
         labs(x="resolution",y="number of clusters") +
         scale_x_continuous(breaks=seq(0,10,1)) +
         scale_y_continuous(breaks=seq(0,10000,50)) +
         theme_jb(),
       width=5, height=5)

umapharmony.SCTccmito.clusPCAharmonyres <- list()
for (res in c(seq(1,4,1))) {
  umapharmony.SCTccmito.clusPCAharmonyres[[as.character(res)]] <- DimPlot(SCC,
                                                                          reduction = "humap_SCTccmito",
                                                                          pt.size = 1,
                                                                          label = T,
                                                                          group.by = paste0("seurat_clusters_all_SCTccmito_PCA_harmony_res",res)) + 
    theme_jb_nogrid()
  ggsave(filename = paste0(OUTPUT,"figs/alltumor-SCTccmito-UMAPharmony-clusPCAharmony-res",res,"-noleg.png"),umapharmony.SCTccmito.clusPCAharmonyres[[as.character(res)]] + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
}
```

## Save 

```{r}
#Single Seurat object
save(file = paste0(OUTPUT,"RData/alltumor.RData"),SCC)
#SessionInfo
writeLines(capture.output(sessionInfo()), paste0(OUTPUT,"RData/alltumor-sessionInfo.txt"))
```
