---
title: "Downstream - Subsetting"
author: "Kai"
date: "10th May 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(ggplot2)
library(ggrepel)

source(paste0("~/tscr/kai/projects/01_SCC_scRNA/results/functions.R"))
OUTPUT = "~/tscr/kai/projects/01_SCC_scRNA/results/pubdatasets/kuertendata/HPVneg_CD45n_analysis/"
```

# Load Seurat object

```{r}
load(paste0(OUTPUT,"RData/HNSCC-fibro.RData"))
```

# Subset cells and evaluate subsetting

```{r}
HNSCC.f.temp <- HNSCC.f
HNSCC.f <- subset(HNSCC.f.temp,cells = names(HNSCC.f.temp$CellType_fibro_SCTccmito_harmony)[HNSCC.f.temp$CellType_fibro_SCTccmito_harmony=="Fibroblasts"])

DefaultAssay(HNSCC.f) <- "RNA"
HNSCC.f@assays$SCT <- NULL
HNSCC.f@assays$SCTccmito <- NULL
for (x in names(HNSCC.f@reductions)) {HNSCC.f@reductions[[x]] <- NULL}
```

# Normalization

## Normalization of RNA assay

```{r SCC_normalization}
#RNA normalization without regressing, necessary for example for differential gene expression not based on SCT.
DefaultAssay(HNSCC.f) <- "RNA"
#log-normalization for slot "data"
HNSCC.f <- NormalizeData(HNSCC.f, 
                         normalization.method = "LogNormalize", #default
                         scale.factor = 10000, #default, scale factor 10000 is recommendation from seurat tutorial
                         assay = "RNA",
                         margin = 1 # default; normalizes across features
) 

#Find Variable Features
HNSCC.f <- FindVariableFeatures(HNSCC.f, 
                                assay = "RNA",
                                selection.method = "vst", #default
                                nfeatures = 2000 #default; only 2000 , 3000 for SCT promises since to "restore" some more VariableFeatures
)

#scaling for slot "scale.data"
HNSCC.f <- ScaleData(HNSCC.f, 
                     features = rownames(GetAssayData(HNSCC.f,assay = "RNA")), 
                     do.scale = T,
                     do.center = T,
                     scale.max = 10,
                     assay = "RNA")
```

## Determine cell cycle

```{r}
HNSCC.f <- CellCycleScoring(HNSCC.f,
                            s.features = cc.genes.updated.2019$s.genes,
                            g2m.features = cc.genes.updated.2019$g2m.genes, 
                            assay="RNA")
```

# Regression of cell cycle and mitochondrial percentage

## Normalization with SCTransform

```{r HNSCC.f_normalization}
HNSCC.f <- SCTransform(HNSCC.f, 
                       ncells = ncol(HNSCC.f),
                       assay="RNA", #default
                       new.assay.name = "SCTccmito", #default; overwrites old SCT
                       do.correct.umi = T, #default change counts slot to corrected counts in new assay
                       variable.features.n = 3000, #default set variable features
                       vars.to.regress = c("S.Score","G2M.Score","percent.mito"), #default optional variables to regress out
                       do.scale = F, #default scale pearson residuals in scale.data slot to have unit variance
                       do.center = T, #default center pearson residuals in scale.data to have mean zero, needed for PCA
                       return.only.var.genes = T, #default scale.data.matrices output assay only contains variable genes
                       seed.use = 1448145)
```

# Dimension reduction

## PCA

```{r}
HNSCC.f <- RunPCA(HNSCC.f, 
                  npcs = 50,  #number of PCs to use
                  assay = "SCTccmito",
                  rev.pca = F, # default, Run cell x gene matrix
                  weight.by.var = T, #default, Weight cell embedding by variance of each PC!)
                  approx = T, #if TRUE, uses irlba instead of prcomp, which is a single value decomposition to approximate PCs, saving computation time (by not calculating all PCs; important for big datasets) and being similar/same to PCs
                  features = VariableFeatures(HNSCC.f,assay = "SCTccmito"), #default
                  reduction.name = "pca_SCTccmito",
                  verbose = F)
```

```{r}
#calculate the % variance explained
PCAvar.SCTccmito <- HNSCC.f@reductions$pca_SCTccmito@stdev^2 / sum(matrixStats::rowVars(as.matrix(HNSCC.f@assays$SCTccmito@scale.data)[VariableFeatures(HNSCC.f,assay = "SCTccmito"),]))

#PCA plot
PCAplot.SCTccmito <- DimPlot(HNSCC.f, reduction = "pca_SCTccmito", dims = c(1,2), group.by = "patients") + 
  labs(x=paste0("principal component 1 (",signif(PCAvar.SCTccmito[1]*100,digits = 2),"% of variance)"),
       y=paste0("principal component 2 (",signif(PCAvar.SCTccmito[2]*100,digits = 2),"% of variance)")) +
  theme_jb_nogrid()

#elbow plot
PCAelbowplot.SCTccmito <- ggplot(data=data.frame(var=PCAvar.SCTccmito,
                                                 PC=1:length(PCAvar.SCTccmito))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=var)) +
  labs(x="principal component", y="% of variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

# cumulative variance Plot
PCAsumplot.SCTccmito <- ggplot(data=data.frame(cumvar=cumsum(PCAvar.SCTccmito),PC=1:length(PCAvar.SCTccmito))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=cumvar)) +
  labs(x="principal component", y="% of cumulative variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-PCA.svg"),PCAplot.SCTccmito, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-PCA-noleg.svg"),PCAplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-PCA-elbow.svg"),PCAelbowplot.SCTccmito, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-PCA-elbow-noleg.svg"),PCAelbowplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-PCA-variancesum.svg"),PCAsumplot.SCTccmito, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-PCA-variancesum-noleg.svg"),PCAsumplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)

PCAelbowplot.SCTccmito
PCAsumplot.SCTccmito
PCAplot.SCTccmito
```

## Graph-based clustering

```{r FindClusters}
numberofPCs = 20

DefaultAssay(HNSCC.f) <- "SCTccmito"
#based on harmony
HNSCC.f <- FindNeighbors(HNSCC.f,
                         dims = 1:numberofPCs,
                         reduction = "pca_SCTccmito",
                         assay = "SCTccmito")
HNSCC.f <- FindClusters(HNSCC.f,
                        resolution = 2,
                        random.seed=100)

HNSCC.f$seurat_clusters_fibrofilt_SCTccmito_PCA <- HNSCC.f$seurat_clusters
HNSCC.f$seurat_clusters <- NULL
```

## UMAP

```{r UMAP}
HNSCC.f <- RunUMAP(HNSCC.f,
                   dims = 1:numberofPCs,
                   assay = "SCTccmito",
                   umap.method = "uwot", # Seurat default
                   graph=NULL, #default
                   reduction="pca_SCTccmito",
                   reduction.name = "umap_SCTccmito",
                   reduction.key = "umap_SCTccmito"
)
```

```{r}
UMAP.SCTccmito.clusPCA <- DimPlot(HNSCC.f, 
                                  reduction = "umap_SCTccmito", 
                                  pt.size = 1, 
                                  label = T, 
                                  group.by = "seurat_clusters_fibrofilt_SCTccmito_PCA") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAP.SCTccmito.patients <- DimPlot(HNSCC.f, 
                                   reduction = "umap_SCTccmito", 
                                   pt.size = 1, 
                                   label = F, 
                                   group.by = "patients") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid() 

UMAP.SCTccmito.Phase <- DimPlot(HNSCC.f, 
                                reduction = "umap_SCTccmito", 
                                pt.size = 1, 
                                label = F, 
                                group.by = "Phase") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()
UMAP.SCTccmito.infstatus <- DimPlot(HNSCC.f, 
                                    reduction = "umap_SCTccmito", 
                                    pt.size = 1, 
                                    label = F, 
                                    group.by = "Inflammation.status") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAP.SCTccmito.QC <- FeaturePlot(HNSCC.f, 
                                 features = c("nCount_RNA_log10", "nFeature_RNA_log10", "percent.mito", "n.exp.hkgenes","S.Score","G2M.Score"), 
                                 reduction = "umap_SCTccmito", 
                                 cols = c("grey95", "blue"),
                                 pt.size = 1,
                                 ncol = 2,
                                 order = T,
                                 slot = "data")

ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAP-clusPCA-noleg.svg"),UMAP.SCTccmito.clusPCA + theme(legend.position = "none"),width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAP-patients.svg"),UMAP.SCTccmito.patients,width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAP-patients-noleg.svg"),UMAP.SCTccmito.patients + theme(legend.position = "none"),width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAP-Phase.svg"),UMAP.SCTccmito.Phase,width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAP-Phase-noleg.svg"),UMAP.SCTccmito.Phase + theme(legend.position = "none"),width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAP-QC.svg"),UMAP.SCTccmito.QC, width=10, height=15)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAP-infstatus.svg"),UMAP.SCTccmito.infstatus,width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAP-infstatus-noleg.svg"),UMAP.SCTccmito.infstatus + theme(legend.position = "none"),width=5,height=5)

UMAP.SCTccmito.clusPCA
UMAP.SCTccmito.patients
UMAP.SCTccmito.Phase
UMAP.SCTccmito.QC
UMAP.SCTccmito.infstatus
```

# Harmony batch correction

```{r}
library(harmony)
set.seed(42)
HNSCC.f <- RunHarmony(HNSCC.f,
                      reduction="pca_SCTccmito",
                      group.by.vars = "patients",
                      reduction.save = "harmony_SCTccmito",
                      plot_convergence = T,
                      max.iter.harmony = 30,
                      assay.use = "SCTccmito",
                      project.dim = F)
```

## Graph-based clustering

```{r FindClusters}
numberofPCs = 20

DefaultAssay(HNSCC.f) <- "SCTccmito"
#based on harmony
HNSCC.f <- FindNeighbors(HNSCC.f,
                         dims = 1:numberofPCs,
                         reduction = "harmony_SCTccmito",
                         assay = "SCTccmito")
HNSCC.f <- FindClusters(HNSCC.f,
                        resolution = 0.8,
                        random.seed=100)

HNSCC.f$seurat_clusters_fibrofilt_SCTccmito_PCA_harmony <- HNSCC.f$seurat_clusters
HNSCC.f$seurat_clusters <- NULL
```

## UMAP

```{r UMAP}
HNSCC.f <- RunUMAP(HNSCC.f,
                   dims = 1:numberofPCs,
                   assay = "SCTccmito",
                   umap.method = "uwot", # Seurat default
                   graph=NULL, #default
                   reduction="harmony_SCTccmito",
                   reduction.name = "humap_SCTccmito",
                   reduction.key = "humap_SCTccmito"
)
```

```{r}
UMAPharmony.SCTccmito.clusPCA <- DimPlot(HNSCC.f, 
                                         reduction = "humap_SCTccmito", 
                                         pt.size = 1, 
                                         label = T, 
                                         group.by = "seurat_clusters_fibrofilt_SCTccmito_PCA_harmony") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAPharmony.SCTccmito.patients <- DimPlot(HNSCC.f, 
                                          reduction = "humap_SCTccmito", 
                                          pt.size = 1, 
                                          label = F, 
                                          group.by = "patients") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid() 

UMAPharmony.SCTccmito.Phase <- DimPlot(HNSCC.f, 
                                       reduction = "humap_SCTccmito", 
                                       pt.size = 1, 
                                       label = F, 
                                       group.by = "Phase") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAPharmony.SCTccmito.infstatus <- DimPlot(HNSCC.f, 
                                           reduction = "humap_SCTccmito", 
                                           pt.size = 1, 
                                           label = F, 
                                           group.by = "Inflammation.status") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAPharmony.SCTccmito.QC <- FeaturePlot(HNSCC.f, 
                                        features = c("nCount_RNA_log10", "nFeature_RNA_log10", "percent.mito", "n.exp.hkgenes","S.Score","G2M.Score"), 
                                        reduction = "humap_SCTccmito", 
                                        cols = c("grey95", "blue"),
                                        pt.size = 1,
                                        ncol = 2,
                                        order = T,
                                        slot = "data")

ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-clusPCA-noleg.svg"),UMAPharmony.SCTccmito.clusPCA + theme(legend.position = "none"),width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-patients.svg"),UMAPharmony.SCTccmito.patients,width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-patients-noleg.svg"),UMAPharmony.SCTccmito.patients + theme(legend.position = "none"),width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-Phase.svg"),UMAPharmony.SCTccmito.Phase,width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-Phase-noleg.svg"),UMAPharmony.SCTccmito.Phase + theme(legend.position = "none"),width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-QC.svg"),UMAPharmony.SCTccmito.QC, width=10, height=15)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-infstatus.svg"),UMAPharmony.SCTccmito.infstatus,width=5,height=5)
ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-infstatus-noleg.svg"),UMAPharmony.SCTccmito.infstatus + theme(legend.position = "none"),width=5,height=5)

UMAPharmony.SCTccmito.clusPCA
UMAPharmony.SCTccmito.patients
UMAPharmony.SCTccmito.Phase
UMAPharmony.SCTccmito.QC
UMAPharmony.SCTccmito.infstatus
```

## Graph-based clustering at different resolutions

```{r FindClusters}
DefaultAssay(HNSCC.f) <- "SCTccmito"
numberofPCs = 20
#based on harmony
for (res in c(seq(1,4,1))) {
  HNSCC.f <- FindNeighbors(HNSCC.f,
                           dims = 1:numberofPCs,
                           reduction = "harmony_SCTccmito",
                           assay = "SCTccmito")
  HNSCC.f <- FindClusters(HNSCC.f,
                          resolution = res,
                          random.seed=100)
  
  HNSCC.f[[paste0("seurat_clusters_fibrofilt_SCTccmito_PCA_harmony_res",res)]] <- HNSCC.f$seurat_clusters
  HNSCC.f$seurat_clusters <- NULL
}

ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-clusPCAharmony-overview.svg"),
       ggplot() + geom_bar(stat="identity", 
                           mapping = aes(y=unlist(lapply(seq(1,4,1),function(res) length(table(HNSCC.f[[paste0("seurat_clusters_fibrofilt_SCTccmito_PCA_harmony_res",res)]])))),x=seq(1,4,1))) +
         labs(x="resolution",y="number of clusters") +
         scale_y_continuous(breaks=seq(0,100,1)) +
         scale_x_continuous(breaks=seq(0,5,1)) +
         theme_jb(),
       width=5, height=5)

ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-clusPCAharmony-overviewcells.svg"),
       ggplot() + geom_bar(stat="identity", 
                           mapping = aes(y=unlist(lapply(seq(1,4,1),function(res) mean(table(HNSCC.f[[paste0("seurat_clusters_fibrofilt_SCTccmito_PCA_harmony_res",res)]])))),x=seq(1,4,1))) +
         labs(x="resolution",y="number of clusters") +
         scale_x_continuous(breaks=seq(0,10,1)) +
         scale_y_continuous(breaks=seq(0,10000,50)) +
         theme_jb(),
       width=5, height=5)

umapharmony.SCTccmito.clusPCAharmonyres <- list()
for (res in c(seq(1,4,1))) {
  umapharmony.SCTccmito.clusPCAharmonyres[[as.character(res)]] <- DimPlot(HNSCC.f,
                                                                          reduction = "humap_SCTccmito",
                                                                          pt.size = 1,
                                                                          label = T,
                                                                          group.by = paste0("seurat_clusters_fibrofilt_SCTccmito_PCA_harmony_res",res)) + 
    theme_jb_nogrid()
  ggsave(filename = paste0(OUTPUT,"figs/HNSCC-fibrofilt-SCTccmito-UMAPharmony-clusPCAharmony-res",res,"-noleg.png"),umapharmony.SCTccmito.clusPCAharmonyres[[as.character(res)]] + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
}
```

#Save

```{r}
#Single Seurat object
save(file = paste0(OUTPUT,"RData/HNSCC-fibrofilt.RData"),HNSCC.f)
#SessionInfo
writeLines(capture.output(sessionInfo()), paste0(OUTPUT,"RData/HNSCC-fibrofilt-sessionInfo.txt"))
```