---
title: "Downstream - Subsetting"
author: "Kai"
date: "10th May 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(ggplot2)
library(ggrepel)
library(ComplexHeatmap)
library(RColorBrewer)
library(SingleR)
library(fgsea)
library(cowplot)

source(paste0("~/tscr/kai/projects/01_SCC_scRNA/results/functions.R"))
OUTPUT = "~/tscr/kai/projects/01_SCC_scRNA/results/samples/cohort/all/rp/"
```

# Load Seurat object

```{r}
load(paste0(OUTPUT,"RData/SCCfilt.RData"))
```

# Subset cells

```{r}
SCC.my <- subset(SCC,cells = names(SCC$CellType_SCTccmito_harmony[SCC$CellType_SCTccmito_harmony %in% c("Myeloid cells","pDCs","RBCs")])) 
rm(SCC)

DefaultAssay(SCC.my) <- "RNA"
SCC.my@assays$SCT <- NULL
SCC.my@assays$SCTccmito <- NULL
for (x in names(SCC.my@reductions)) {SCC.my@reductions[[x]] <- NULL}
```

# Normalization

## Normalization of RNA assay

```{r SCC_normalization}
#RNA normalization without regressing, necessary for example for differential gene expression not based on SCT.
DefaultAssay(SCC.my) <- "RNA"
#log-normalization for slot "data"
SCC.my <- NormalizeData(SCC.my, 
                        normalization.method = "LogNormalize", #default
                        scale.factor = 10000, #default, scale factor 10000 is recommendation from seurat tutorial
                        assay = "RNA",
                        margin = 1 # default; normalizes across features
) 

#Find Variable Features
SCC.my <- FindVariableFeatures(SCC.my, 
                               assay = "RNA",
                               selection.method = "vst", #default
                               nfeatures = 2000 #default; only 2000 , 3000 for SCT promises since to "restore" some more VariableFeatures
)

#scaling for slot "scale.data"
SCC.my <- ScaleData(SCC.my, 
                    features = VariableFeatures(SCC.my,assay = "RNA"), 
                    do.scale = T,
                    do.center = T,
                    scale.max = 10,
                    assay = "RNA")
```

## Determine cell cycle

```{r}
SCC.my <- CellCycleScoring(SCC.my,
                           s.features = cc.genes.updated.2019$s.genes,
                           g2m.features = cc.genes.updated.2019$g2m.genes, 
                           assay="RNA")
```

# Cell Cycle and mitochondrial percentage regression

## Normalization with SCTransform

```{r SCC_normalization}
SCC.my <- SCTransform(SCC.my, 
                      ncells = ncol(SCC.my),
                      assay="RNA", #default
                      new.assay.name = "SCTccmito", #default; overwrites old SCTccmito
                      do.correct.umi = T, #default change counts slot to corrected counts in new assay
                      variable.features.n = 3000, #default set variable features
                      vars.to.regress = c("S.Score","G2M.Score","percent.mito"), #default optional variables to regress out
                      do.scale = F, #default scale pearson residuals in scale.data slot to have unit variance
                      do.center = T, #default center pearson residuals in scale.data to have mean zero, needed for PCA
                      return.only.var.genes = T, #default scale.data.matrices output assay only contains variable genes
                      seed.use = 1448145)
```

# Dimension reduction

## PCA

always based on SCTccmito assay

```{r}
SCC.my <- RunPCA(SCC.my, 
                 npcs = 50,  #number of PCs to use
                 assay = "SCTccmito",
                 rev.pca = F, # default, Run cell x gene matrix
                 weight.by.var = T, #default, Weight cell embedding by variance of each PC!)
                 approx = T, #if TRUE, uses irlba instead of prcomp, which is a single value decomposition to approximate PCs, saving computation time (by not calculating all PCs; important for big datasets) and being similar/same to PCs
                 features = VariableFeatures(SCC.my,assay = "SCTccmito"), #default
                 reduction.name = "pca_SCTccmito",
                 verbose = F)
```

```{r}
#calculate the % variance explained
PCAvar.SCTccmito <- SCC.my@reductions$pca_SCTccmito@stdev^2 / sum(matrixStats::rowVars(as.matrix(SCC.my@assays$SCTccmito@scale.data)[VariableFeatures(SCC.my,assay = "SCTccmito"),]))

#PCA plot
PCAplot.SCTccmito <- DimPlot(SCC.my, reduction = "pca_SCTccmito", dims = c(1,2), group.by = "hash.ID") + 
  labs(x=paste0("principal component 1 (",signif(PCAvar.SCTccmito[1]*100,digits = 2),"% variance)"),
       y=paste0("principal component 2 (",signif(PCAvar.SCTccmito[2]*100,digits = 2),"% variance)")) +
  theme_jb()

#elbow plot
PCAelbowplot.SCTccmito <- ggplot(data=data.frame(var=PCAvar.SCTccmito,
                                                 PC=1:length(PCAvar.SCTccmito))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=var)) +
  labs(x="principal component", y="% variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

# cumulative Variance Plot
PCAsumplot.SCTccmito <- ggplot(data=data.frame(cumvar=cumsum(PCAvar.SCTccmito),PC=1:length(PCAvar.SCTccmito))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=cumvar)) +
  labs(x="principal component", y="% cumulative variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-PCA.svg"),PCAplot.SCTccmito, width = 5, height = 5)
# without legends
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-PCA-noleg.svg"),PCAplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-PCA-elbow.svg"),PCAelbowplot.SCTccmito, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-PCA-elbow-noleg.svg"),PCAelbowplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-PCA-variancesum.svg"),PCAsumplot.SCTccmito, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-PCA-variancesum-noleg.svg"),PCAsumplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL),width = 5, height = 5)

PCAplot.SCTccmito
PCAelbowplot.SCTccmito
PCAsumplot.SCTccmito
```

# Harmony batch correction

```{r}
library(harmony)
set.seed(42)
SCC.my <- RunHarmony(SCC.my,
                     reduction="pca_SCTccmito",
                     group.by.vars = "lib",
                     reduction.save = "harmony_SCTccmito",
                     plot_convergence = T,
                     max.iter.harmony = 30,
                     assay.use = "SCTccmito",
                     project.dim = F)
```

### Graph-based clustering

```{r FindClusters}
DefaultAssay(SCC.my) <- "SCTccmito"
numberofPCs = 20
#based on harmony
SCC.my <- FindNeighbors(SCC.my,
                        dims = 1:numberofPCs,
                        reduction = "harmony_SCTccmito",
                        assay = "SCTccmito")
SCC.my <- FindClusters(SCC.my,
                       resolution = 0.8,
                       random.seed=100)

SCC.my$seurat_clusters_my_SCTccmito_PCA_harmony <- SCC.my$seurat_clusters
SCC.my$seurat_clusters <- NULL
```

### UMAP

```{r}
SCC.my <- RunUMAP(SCC.my,
                  dims = 1:numberofPCs,
                  assay = "SCTccmito",
                  umap.method = "uwot", # Seurat default
                  graph=NULL, #default
                  reduction="harmony_SCTccmito",
                  reduction.name = "humap_SCTccmito",
                  reduction.key = "humap_SCTccmito"
)
```

```{r}
umapharmony.SCTccmito.clusPCAharmony <- DimPlot(SCC.my,
                                                reduction = "humap_SCTccmito",
                                                pt.size = 1,
                                                label = T,
                                                group.by = "seurat_clusters_my_SCTccmito_PCA_harmony") +
  theme_jb_nogrid()

umapharmony.SCTccmito.celltype <- DimPlot(SCC.my,
                                          reduction = "humap_SCTccmito",
                                          pt.size = 1,
                                          label = F,
                                          group.by = "CellType_SCTccmito_harmony") + 
  theme_jb_nogrid()

umapharmony.SCTccmito.patients <- DimPlot(SCC.my,
                                          reduction = "humap_SCTccmito",
                                          pt.size = 1,
                                          label = F,
                                          group.by = "patient") + 
  theme_jb_nogrid()

umapharmony.SCTccmito.sample <- DimPlot(SCC.my,
                                        reduction = "humap_SCTccmito",
                                        pt.size = 1,
                                        label = F,
                                        group.by = "sample") + 
  theme_jb_nogrid()

umapharmony.SCTccmito.Lok <- DimPlot(SCC.my,
                                     reduction = "humap_SCTccmito",
                                     pt.size = 1,
                                     label = F,
                                     group.by = "Lok") + 
  theme_jb_nogrid()

umapharmony.SCTccmito.lib <- DimPlot(SCC.my,
                                     reduction = "humap_SCTccmito",
                                     pt.size = 1,
                                     label = F,
                                     group.by = "lib") + 
  theme_jb_nogrid()

umapharmony.SCTccmito.Phase <- DimPlot(SCC.my,
                                       reduction = "humap_SCTccmito",
                                       pt.size = 1,
                                       label = F,
                                       group.by = "Phase") + 
  theme_jb_nogrid()

UMAPharmony.SCTccmito.qc <- FeaturePlot(SCC.my, 
                                        features = c("nCount_RNA_log10", "nFeature_RNA_log10", "percent.mito", "n.exp.hkgenes","S.Score","G2M.Score"), 
                                        reduction = "humap_SCTccmito", 
                                        cols = c("grey95", "blue"),
                                        pt.size = 1,
                                        ncol = 2,
                                        order = T,
                                        slot = "data") +
  theme_jb_nogrid()

ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-clusPCAharmony.svg"),umapharmony.SCTccmito.clusPCAharmony)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-clusPCAharmony-noleg.svg"),umapharmony.SCTccmito.clusPCAharmony + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-CelltypeSCTccmitoharmony.svg"),umapharmony.SCTccmito.celltype)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-CelltypeSCTccmitoharmony-noleg.svg"),umapharmony.SCTccmito.celltype + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-patients.svg"),umapharmony.SCTccmito.patients)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-patients-noleg.svg"),umapharmony.SCTccmito.patients + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-sample.svg"),umapharmony.SCTccmito.sample)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-sample-noleg.svg"),umapharmony.SCTccmito.sample + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-Lok.svg"),umapharmony.SCTccmito.Lok)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-Lok-noleg.svg"),umapharmony.SCTccmito.Lok + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-lib.svg"),umapharmony.SCTccmito.lib)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-lib-noleg.svg"),umapharmony.SCTccmito.lib + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-Phase.svg"),umapharmony.SCTccmito.Phase)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-Phase-noleg.svg"),umapharmony.SCTccmito.Phase + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-QC.svg"),UMAPharmony.SCTccmito.qc, width=10, height=10)

umapharmony.SCTccmito.clusPCAharmony
umapharmony.SCTccmito.celltype
umapharmony.SCTccmito.patients
umapharmony.SCTccmito.sample
umapharmony.SCTccmito.Lok
umapharmony.SCTccmito.lib
umapharmony.SCTccmito.Phase
UMAPharmony.SCTccmito.qc
```

# A bit more detailed clustering for better filtering

```{r}
DefaultAssay(SCC.my) <- "SCTccmito"
numberofPCs = 20
#based on harmony
SCC.my <- FindNeighbors(SCC.my,
                        dims = 1:numberofPCs,
                        reduction = "harmony_SCTccmito",
                        assay = "SCTccmito")
SCC.my <- FindClusters(SCC.my,
                       resolution = 4,
                       random.seed=100)

SCC.my$seurat_clusters_my_SCTccmito_PCA_harmony_res4 <- SCC.my$seurat_clusters
SCC.my$seurat_clusters <- NULL

umapharmony.SCTccmito.clusPCAres4 <- DimPlot(SCC.my,
                                             reduction = "humap_SCTccmito",
                                             pt.size = 1,
                                             label = T,
                                             group.by = "seurat_clusters_my_SCTccmito_PCA_harmony_res4") + 
  theme_jb_nogrid()
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-clusPCA-res4-noleg.svg"),umapharmony.SCTccmito.clusPCAres4 + theme(legend.position = "none"), height=5, width = 5)
umapharmony.SCTccmito.clusPCAres4
```

# Marker Gene Expression

```{r}
UMAPharmony.SCTccmito.marker <- lapply(names(celltypemarkers), function(x) FeaturePlot_markers_comb(SCC.my,celltypemarkers[[x]],reduction = "humap_SCTccmito"))
names(UMAPharmony.SCTccmito.marker) <- names(celltypemarkers)

lapply(names(UMAPharmony.SCTccmito.marker), function(x) ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-marker-",x,".png"),UMAPharmony.SCTccmito.marker[[x]],dpi=300, width=ifelse(length(celltypemarkers[[x]])>1,10,5), height=5*ceiling(length(celltypemarkers[[x]])/2)))

UMAPharmony.SCTccmito.marker
```

#SingleR

# clustering

```{r}
DefaultAssay(SCC.my) <- "SCTccmito"
numberofPCs <- 20
#resolution 100
#clustering
SCC.my <- FindNeighbors(SCC.my,
                        dims = 1:numberofPCs,
                        reduction = "harmony_SCTccmito",
                        assay = "SCTccmito")
SCC.my <- FindClusters(SCC.my,
                       resolution = 100,
                       random.seed = 100)

SCC.my$seurat_clusters_my_SCTccmito_PCA_harmony_res100 <- SCC.my$seurat_clusters
SCC.my$seurat_clusters <- NULL

# plot
umapharmony.SCTccmito.clusPCAharmonyres100 <- DimPlot(SCC.my,
                                                      reduction = "humap_SCTccmito",
                                                      pt.size = 1,
                                                      label = F,
                                                      group.by = "seurat_clusters_my_SCTccmito_PCA_harmony_res100") + 
  labs(title=NULL) +
  theme_jb_nogrid() + 
  theme(legend.position = "none")

ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-clusPCAharmony-res100-noleg.png"),umapharmony.SCTccmito.clusPCAharmonyres100, width = 5, height = 5)

umapharmony.SCTccmito.clusPCAharmonyres100
```

## SingleR for high resolution cluster

```{r}
# Load Reference set
monaco.ref <- MonacoImmuneData()

# run singleR
predclus.monaco <- SingleR(test = GetAssayData(SCC.my, slot = "data",assay = "SCTccmito"), 
                           ref = monaco.ref, 
                           assay.type.test = 1,
                           clusters = SCC.my$seurat_clusters_my_SCTccmito_PCA_harmony_res100,
                           labels = monaco.ref$label.fine)

# add singleR annotations to the seurat object
cluslabels.monaco <- predclus.monaco$labels
names(cluslabels.monaco) <- rownames(predclus.monaco)

SCC.my$SingleR.monaco.res100 <- cluslabels.monaco[match(SCC.my$seurat_clusters_my_SCTccmito_PCA_harmony_res100,names(cluslabels.monaco))]

#plot it
UMAPharmony.SCTccmito.singleR_monaco_res100 <- DimPlot(SCC.my, 
                                                       reduction = "humap_SCTccmito", 
                                                       pt.size = 1, 
                                                       label = T, 
                                                       group.by = "SingleR.monaco.res100") +  
  theme_jb_nogrid()
UMAPharmony.SCTccmito.singleR_monaco_res100.nol <- DimPlot(SCC.my, 
                                                           reduction = "humap_SCTccmito", 
                                                           pt.size = 1, 
                                                           label = F, 
                                                           group.by = "SingleR.monaco.res100") +  
  theme_jb_nogrid()

ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-SingleR-monaco_res100.svg"),UMAPharmony.SCTccmito.singleR_monaco_res100, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-SingleR-monaco_res100-noleg.svg"),UMAPharmony.SCTccmito.singleR_monaco_res100 + theme(legend.position = "none"), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-SingleR-monaco_res100-nolabelleg.svg"),UMAPharmony.SCTccmito.singleR_monaco_res100.nol + theme(legend.position = "none"), width = 5, height = 5)

UMAPharmony.SCTccmito.singleR_monaco_res100
```

```{r}
#all types separately
UMAPharmony.SCTccmito.singleR_monaco_res100.type <- list()
for (type in unique(SCC.my$SingleR.monaco.res100)) {
  SCC.my <- AddMetaData(SCC.my, SCC.my$SingleR.monaco.res100 == type, col.name = paste0("SingleR.monaco.res100.",type))
  
  UMAPharmony.SCTccmito.singleR_monaco_res100.type[[type]] <- DimPlot(SCC.my, 
                                                                      reduction = "humap_SCTccmito", 
                                                                      pt.size = 1, 
                                                                      label = F, 
                                                                      order = "TRUE",
                                                                      cols = c("grey95","blue"),
                                                                      group.by = paste0("SingleR.monaco.res100.",gsub("-",".",gsub(" ",".",gsub("/",".",type))))) +  
    ggtitle(type) +
    theme_jb_nogrid()
  ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-SingleR-monaco-res100-",gsub("-",".",gsub(" ",".",gsub("/",".",type))),"-noleg.png"),UMAPharmony.SCTccmito.singleR_monaco_res100.type[[type]] + theme(legend.position = "none"), width = 5, height = 5)
}

UMAPharmony.SCTccmito.singleR_monaco_res100.type
```

# Differential expression

```{r}
SCC.my.clusharmony40 <- calculate_foldchanges(SCC.my,SCC.my$seurat_clusters_my_SCTccmito_PCA_harmony_res4)
#filter for genes with low foldchanges and pct value. This improves GSEA analysis because genes suffering from high dropout (i.e., lowly expressed and low signal) are not considered for ranking, while most up/downregulated genes are still preserved. These are also default values for "FindAllMarkers" function
SCC.my.clusharmony40 <- SCC.my.clusharmony40[(SCC.my.clusharmony40$pct.1 > 0.1 | SCC.my.clusharmony40$pct.2 > 0.1) & abs(SCC.my.clusharmony40$avg_log2FC)>0.25,]
```

## Volcano Plots 

```{r}
#set thres function for selecting genes to mark in volcano plot
logFCcutoff_clusharmony40 = function(x) {
  c(sort(SCC.my.clusharmony40[SCC.my.clusharmony40$cluster==x,]$avg_log2FC, decreasing = T)[11])
}
pctdiffcutoff_clusharmony40 = function(x) {
  c(sort(SCC.my.clusharmony40[SCC.my.clusharmony40$cluster==x,]$pct.1 - SCC.my.clusharmony40[SCC.my.clusharmony40$cluster==x,]$pct.2, decreasing = T)[21])
}

#create volcano plots with pct difference on y-axis
SCC.my.clusharmony40.volpct <- lapply(levels(SCC.my.clusharmony40$cluster), function(x) ggvolcano(SCC.my.clusharmony40[SCC.my.clusharmony40$cluster==x,],logFCcutoff_clusharmony40(x),onlypos = T,pctdiff = F) + ggtitle(x))
names(SCC.my.clusharmony40.volpct) <- levels(SCC.my.clusharmony40$cluster)

SCC.my.clusharmony40.volpctdiff <- lapply(levels(SCC.my.clusharmony40$cluster), function(x) ggvolcano(SCC.my.clusharmony40[SCC.my.clusharmony40$cluster==x,],pctdiffcutoff_clusharmony40(x),onlypos = T,pctdiff = T) + ggtitle(x))
names(SCC.my.clusharmony40.volpctdiff) <- levels(SCC.my.clusharmony40$cluster)

#save volcano plots
for (i in names(SCC.my.clusharmony40.volpct)) {
  ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-diffexpr-SCTccmito-clusharmony40-volcanopct-",gsub("[/]","",sub(" ","",sub("[)]","",sub("[(]","",i)))),".svg"),SCC.my.clusharmony40.volpct[[i]])
}
for (i in names(SCC.my.clusharmony40.volpctdiff)) {
  ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-diffexpr-SCTccmito-clusharmony40-volcanopctdiff-",gsub("[/]","",sub(" ","",sub("[)]","",sub("[(]","",i)))),".svg"),SCC.my.clusharmony40.volpctdiff[[i]])
}

SCC.my.clusharmony40.volpct
```

## Gene Set Enrichment Analysis (GSEA)

```{r}
FCranks.clusharmony40 <- list()
for (cl in unique(SCC.my.clusharmony40$cluster)) {
  #rank DGE results by foldchange
  FCranks.clusharmony40[[cl]] <- SCC.my.clusharmony40[SCC.my.clusharmony40$cluster==cl,]$avg_log2FC
  names(FCranks.clusharmony40[[cl]]) <- SCC.my.clusharmony40[SCC.my.clusharmony40$cluster==cl,]$gene
  
  #convert -Inf to -1000
  FCranks.clusharmony40[[cl]][is.infinite(FCranks.clusharmony40[[cl]])] <- -1000
}

#read in MsigDB pathway from downloads, in this case GO: BP
GOBP <- gmtPathways("~/tscr/kai/db/msigdb/c5.bp.v7.1.symbols.gmt")
HM <- gmtPathways("~/tscr/kai/db/msigdb/h.all.v7.1.symbols.gmt")

GS <- c(GOBP,HM)

#run fgsea
GS.gsea.clusharmony40 <- list()
for (cl in levels(SCC.my.clusharmony40$cluster)) {
  GS.gsea.clusharmony40[[cl]] <- fgsea(pathways=GS,
                                       stats=FCranks.clusharmony40[[cl]],
                                       minSize=15,
                                       maxSize=500,
                                       nperm=10000)
}

#remove GO in name and put to lower cases
GS.gsea.clusharmony40.plot <- lapply(GS.gsea.clusharmony40,function(x) {x$pathway <- gsub("_"," ",tolower(sub("GO_","",x$pathway))); x})
#abbreviate some GO terms
GS.gsea.clusharmony40.plot <- lapply(GS.gsea.clusharmony40.plot,function(x) {x$pathway <- gsub("endoplasmic reticulum","ER",
                                                                                               gsub("nonsense mediated decay","NMD",x$pathway)); x})
#Plotting
numberofGS = 5

GS.gseavis.clusharmony40 <- lapply(names(GS.gsea.clusharmony40.plot), function(x) {
  try(GSEAvis(GSEAdf = GS.gsea.clusharmony40.plot[[x]],
              pvalthres = NULL,
              NESthres = NULL,
              npathways = numberofGS,
              sort = "NES",
              positive = T) +
        labs(x=x,y="") +
        scale_fill_gradientn(limits = c(0,-log10(min(unlist(lapply(GS.gsea.clusharmony40.plot,function(x) {y <- x[x$NES>0,]; y[order(y$NES,decreasing = T),]$padj[1:numberofGS]}))))),
                             colours = c("white","blue")) +
        coord_flip(ylim = c(0,max(unlist(lapply(GS.gsea.clusharmony40.plot,function(x) {y <- x[x$NES>0,]; y[order(y$NES,decreasing = T),]$NES[1:numberofGS]}))))) +
        scale_y_continuous(breaks = seq(0,20,0.5))
  )}) 
names(GS.gseavis.clusharmony40) <- names(GS.gsea.clusharmony40.plot)
lapply(names(GS.gsea.clusharmony40.plot), function(x) try(ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-diffexpr-SCTccmito-clusharmony40-GS-top",numberofGS,"byNES-",gsub("[/]","",sub(" ","",sub("[)]","",sub("[(]","",x)))),".svg"),GS.gseavis.clusharmony40[[x]],width=10, height=5)))

#All GSEA Plots in one file
GS.gseavis.clusharmony40.comb <- lapply(1:(length(GS.gseavis.clusharmony40)-1), function(x) {
  GS.gseavis.clusharmony40[[x]] + 
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())
})
GS.gseavis.clusharmony40.comb[[length(GS.gseavis.clusharmony40.comb)+1]] <- GS.gseavis.clusharmony40[[length(GS.gseavis.clusharmony40)]]
try(ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-diffexpr-SCTccmito-clusharmony40-GS-top",numberofGS,"byNES-all.svg"),
           plot_grid(plotlist = GS.gseavis.clusharmony40.comb,ncol=1,align = "v",labels=NULL),
           width=10, 
           height=20))

try(ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-diffexpr-SCTccmito-clusharmony40-GS-top",numberofGS,"byNES-all-noleg.svg"),
           plot_grid(plotlist = lapply(GS.gseavis.clusharmony40.comb,function(x) x + theme(legend.position = "none")),ncol=1,align = "v",labels=NULL),
           width=10, 
           height=20))

GS.gseavis.clusharmony40.comb <- lapply(1:(length(GS.gseavis.clusharmony40)), function(x) {
  GS.gseavis.clusharmony40[[x]] + 
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())
})
try(ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-diffexpr-SCTccmito-clusharmony40-GS-top",numberofGS,"byNES-all-nolegaxis.svg"),
           plot_grid(plotlist = lapply(GS.gseavis.clusharmony40.comb,function(x) x + theme(legend.position = "none")),ncol=1,align = "v",labels=NULL),
           width=10, 
           height=20))
```

## Heatmap of top differentially expressed genes

```{r}
topgenes = 10

#choose those top genes
SCC.my.clusharmony40.percluster <- lapply(unique(SCC.my.clusharmony40$cluster), function(x) SCC.my.clusharmony40[SCC.my.clusharmony40$cluster==x,])
names(SCC.my.clusharmony40.percluster) <- unique(SCC.my.clusharmony40$cluster)

top.logFC.clusharmony40 <- list()
highlightgenes.clusharmony40 <- list()
for (cl in names(SCC.my.clusharmony40.percluster)) {
  #top avg FC
  top.logFC.clusharmony40[[cl]] <- head(SCC.my.clusharmony40.percluster[[cl]][order(SCC.my.clusharmony40.percluster[[cl]]$avg_log2FC,decreasing = T),]$gene, n=topgenes)
  #create dataframe
  highlightgenes.clusharmony40[[cl]] <- data.frame(genes=top.logFC.clusharmony40[[cl]], state=cl)
}

# summarize in one dataframe
highlightgenes.clusharmony40 <- as.data.frame(data.table::rbindlist(highlightgenes.clusharmony40))

SCC.my <- ScaleData(SCC.my,
                    features = as.character(highlightgenes.clusharmony40$genes),
                    do.scale = T,
                    do.center = T,
                    scale.max = 10,
                    assay = "RNA")
expr.RNAscaledata <- as.data.frame(GetAssayData(SCC.my, assay = "RNA", slot = "scale.data"))
#subset expression matrix by highlight genes
expr.RNAscaledata.hl.clusharmony40 <- expr.RNAscaledata[as.character(highlightgenes.clusharmony40$genes),]
#set colorbreaks as above
colorbreaks.RNAscaledata.hl.clusharmony40 <- c(-max(abs(expr.RNAscaledata.hl.clusharmony40)),
                                               -(mean(unlist(abs(expr.RNAscaledata.hl.clusharmony40))[unlist(abs(expr.RNAscaledata.hl.clusharmony40))!=0]) + 2*sd(unlist(abs(expr.RNAscaledata.hl.clusharmony40))[unlist(abs(expr.RNAscaledata.hl.clusharmony40))!=0])),
                                               0,
                                               mean(unlist(abs(expr.RNAscaledata.hl.clusharmony40))[unlist(abs(expr.RNAscaledata.hl.clusharmony40))!=0]) + 2*sd(unlist(abs(expr.RNAscaledata.hl.clusharmony40))[unlist(abs(expr.RNAscaledata.hl.clusharmony40))!=0]),
                                               max(abs(expr.RNAscaledata.hl.clusharmony40)))

#Heatmap clustered by columns and rows, rows split by differential expression of certain cluster
expr.RNAscaledata.hl.clusharmony40.hm <- Heatmap(expr.RNAscaledata.hl.clusharmony40,
                                                 name="scale.data",
                                                 col=circlize::colorRamp2(breaks=colorbreaks.RNAscaledata.hl.clusharmony40, colors = c("red","purple", "black", "yellow","white")),
                                                 show_column_names = F,
                                                 show_row_names = T,
                                                 cluster_columns = T,
                                                 cluster_column_slices = F,
                                                 clustering_method_columns = "ward.D2",
                                                 clustering_distance_columns = "euclidean",
                                                 column_split = SCC.my$seurat_clusters_my_SCTccmito_PCA_harmony_res4,
                                                 row_split = factor(highlightgenes.clusharmony40$state, levels = levels(SCC.my$seurat_clusters_my_SCTccmito_PCA_harmony_res4)),
                                                 cluster_rows = F, row_order = highlightgenes.clusharmony40$genes,
                                                 row_title_rot = 0,
                                                 column_title_rot = 90,
                                                 #graphic options
                                                 row_names_gp = gpar(fontsize=7),
                                                 column_names_gp = gpar(fontsize=11),
                                                 row_title_gp = gpar(fontsize=11),
                                                 column_title_gp = gpar(fontsize=11),
                                                 heatmap_legend_param = list(title_gp = gpar(fontsize=11), labels_gp = gpar(fontsize=11)),
                                                 border = T,
                                                 use_raster = T,
                                                 raster_device = "CairoPNG",
                                                 raster_quality = 4,
                                                 height = unit(20,"in"),
                                                 width = unit(10,"in"))
#Save Heatmap
pdf(paste0(OUTPUT,"figs/SCCfilt-my-diffexpr-SCTccmito-clusharmony40-HM-RNAscaledata-top",as.character(topgenes),".pdf"), width=20, height=30)
expr.RNAscaledata.hl.clusharmony40.hm
dev.off()
```

# Remove non myeloid cells

```{r}
SCC.my$CellType_my <- as.character(SCC.my$seurat_clusters_my_SCTccmito_PCA_harmony_res4)
SCC.my$CellType_my[SCC.my$CellType_my=="17"] <- "B-Cells"
SCC.my$CellType_my[SCC.my$CellType_my=="23"] <- "T-Cells"
SCC.my$CellType_my[SCC.my$CellType_my=="24"] <- "T-Cells"
SCC.my$CellType_my[SCC.my$CellType_my=="28"] <- "Fibroblasts / Tumor"

SCC.my$CellType_my[!SCC.my$CellType_my %in% c("T-Cells","B-Cells","Fibroblasts / Tumor")] <- "myeloid cells"

umapharmony.SCTccmito.celltypemy <- DimPlot(SCC.my,
                                            reduction = "humap_SCTccmito",
                                            pt.size = 1,
                                            label = T,
                                            group.by = "CellType_my") + 
  theme_jb_nogrid()

ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-my-SCTccmito-UMAPharmony-Celltypemy-noleg.svg"),umapharmony.SCTccmito.celltypemy + theme(legend.position = "none"), height=5, width = 5)
umapharmony.SCTccmito.celltypetcells
```

#Save

```{r}
#Single Seurat object
save(file = paste0(OUTPUT,"RData/SCCfilt-my.RData"),SCC.my)
#SessionInfo
writeLines(capture.output(sessionInfo()), paste0(OUTPUT,"RData/SCCfilt-my-sessionInfo.txt"))
```
