---
title: "Downstream - Subsetting"
author: "Kai"
date: "10th May 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Seurat)
library(ggplot2)
library(ggrepel)
library(ComplexHeatmap)
library(RColorBrewer)
library(SingleR)
library(fgsea)
library(cowplot)

source(paste0("~/tscr/kai/projects/01_SCC_scRNA/results/functions.R"))
OUTPUT = "~/tscr/kai/projects/01_SCC_scRNA/results/samples/cohort/all/rp/"
```

# Load Seurat object

```{r}
load(paste0(OUTPUT,"RData/SCCfilt.RData"))
```

# Subset cells

```{r}
SCC.T <- subset(SCC,cells = names(SCC$CellType_SCTccmito_harmony[SCC$CellType_SCTccmito_harmony %in% c("T-Cells")])) 
rm(SCC)

DefaultAssay(SCC.T) <- "RNA"
SCC.T@assays$SCT <- NULL
SCC.T@assays$SCTccmito <- NULL
for (x in names(SCC.T@reductions)) {SCC.T@reductions[[x]] <- NULL}
```

# Normalization

## Normalization of RNA assay

```{r SCC_normalization}
#RNA normalization without regressing, necessary for example for differential gene expression not based on SCT.
DefaultAssay(SCC.T) <- "RNA"
#log-normalization for slot "data"
SCC.T <- NormalizeData(SCC.T, 
                       normalization.method = "LogNormalize", #d-efault
                       scale.factor = 10000, #default, scale factor 10000 is recommendation from seurat tutorial
                       assay = "RNA",
                       margin = 1 # default; normalizes across features
) 

#Find Variable Features
SCC.T <- FindVariableFeatures(SCC.T, 
                              assay = "RNA",
                              selection.method = "vst", #default
                              nfeatures = 2000 #default; only 2000 , 3000 for SCT promises since to "restore" some more VariableFeatures
)

#scaling for slot "scale.data"
SCC.T <- ScaleData(SCC.T, 
                   features = VariableFeatures(SCC.T,assay="RNA"), 
                   do.scale = T,
                   do.center = T,
                   scale.max = 10,
                   assay = "RNA")
```

## Determine cell cycle

```{r}
SCC.T <- CellCycleScoring(SCC.T,
                          s.features = cc.genes.updated.2019$s.genes,
                          g2m.features = cc.genes.updated.2019$g2m.genes, 
                          assay="RNA")
```

# Cell Cycle and mitochondrial percentage regression

## Normalization with SCTransform

More information: https://doi.org/10.1186/s13059-019-1874-1

```{r SCC_normalization}
SCC.T <- SCTransform(SCC.T, 
                     ncells = ncol(SCC.T),
                     assay="RNA", #default
                     new.assay.name = "SCTccmito", #default; overwrites old SCTccmito
                     do.correct.umi = T, #default change counts slot to corrected counts in new assay
                     variable.features.n = 3000, #default set variable features
                     vars.to.regress = c("S.Score","G2M.Score","percent.mito"), #default optional variables to regress out
                     do.scale = F, #default scale pearson residuals in scale.data slot to have unit variance
                     do.center = T, #default center pearson residuals in scale.data to have mean zero, needed for PCA
                     return.only.var.genes = T, #default scale.data.matrices output assay only contains variable genes
                     seed.use = 1448145)
```

# Dimension reduction

## PCA

always based on SCTccmito assay

```{r}
SCC.T <- RunPCA(SCC.T, 
                npcs = 50,  #number of PCs to use
                assay = "SCTccmito",
                rev.pca = F, # default, Run cell x gene matrix
                weight.by.var = T, #default, Weight cell embedding by variance of each PC!)
                approx = T, #if TRUE, uses irlba instead of prcomp, which is a single value decomposition to approximate PCs, saving computation time (by not calculating all PCs; important for big datasets) and being similar/same to PCs
                features = VariableFeatures(SCC.T,assay = "SCTccmito"), #default
                reduction.name = "pca_SCTccmito",
                verbose = F)
```

```{r}
#calculate the % variance explained
PCAvar.SCTccmito <- SCC.T@reductions$pca_SCTccmito@stdev^2 / sum(matrixStats::rowVars(as.matrix(SCC.T@assays$SCTccmito@scale.data)[VariableFeatures(SCC.T,assay = "SCTccmito"),]))

#PCA plot
PCAplot.SCTccmito <- DimPlot(SCC.T, reduction = "pca_SCTccmito", dims = c(1,2), group.by = "hash.ID") + 
  labs(x=paste0("principal component 1 (",signif(PCAvar.SCTccmito[1]*100,digits = 2),"% variance)"),
       y=paste0("principal component 2 (",signif(PCAvar.SCTccmito[2]*100,digits = 2),"% variance)")) +
  theme_jb()

#elbow plot
PCAelbowplot.SCTccmito <- ggplot(data=data.frame(var=PCAvar.SCTccmito,
                                             PC=1:length(PCAvar.SCTccmito))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=var)) +
  labs(x="principal component", y="% variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

# cumulative Variance Plot
PCAsumplot.SCTccmito <- ggplot(data=data.frame(cumvar=cumsum(PCAvar.SCTccmito),PC=1:length(PCAvar.SCTccmito))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=cumvar)) +
  labs(x="principal component", y="% cumulative variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-PCA.svg"),PCAplot.SCTccmito, width = 5, height = 5)
# without legends
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-PCA-noleg.svg"),PCAplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-PCA-elbow.svg"),PCAelbowplot.SCTccmito, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-PCA-elbow-noleg.svg"),PCAelbowplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-PCA-variancesum.svg"),PCAsumplot.SCTccmito, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-PCA-variancesum-noleg.svg"),PCAsumplot.SCTccmito + theme(legend.position = "none") + labs(title = NULL),width = 5, height = 5)

PCAplot.SCTccmito
PCAelbowplot.SCTccmito
PCAsumplot.SCTccmito
```

# Harmony batch correction

```{r}
library(harmony)
set.seed(42)
SCC.T <- RunHarmony(SCC.T,
                    reduction="pca_SCTccmito",
                    group.by.vars = "lib",
                    reduction.save = "harmony_SCTccmito",
                    plot_convergence = T,
                    max.iter.harmony = 30,
                    assay.use = "SCTccmito",
                    project.dim = F)
```

### Graph-based clustering

```{r FindClusters}
DefaultAssay(SCC.T) <- "SCTccmito"
numberofPCs = 20
#based on harmony
SCC.T <- FindNeighbors(SCC.T,
                       dims = 1:numberofPCs,
                       reduction = "harmony_SCTccmito",
                       assay = "SCTccmito")
SCC.T <- FindClusters(SCC.T,
                      resolution = 0.8,
                      random.seed=100)

SCC.T$seurat_clusters_TCells_SCTccmito_PCA_harmony <- SCC.T$seurat_clusters
SCC.T$seurat_clusters <- NULL
```

### UMAP

```{r}
SCC.T <- RunUMAP(SCC.T,
                 dims = 1:numberofPCs,
                 assay = "SCTccmito",
                 umap.method = "uwot", # Seurat default
                 graph=NULL, #default
                 reduction="harmony_SCTccmito",
                 reduction.name = "humap_SCTccmito",
                 reduction.key = "humap_SCTccmito"
)
```

```{r}
umapharmony.SCTccmito.clusPCAharmony <- DimPlot(SCC.T,
                                                reduction = "humap_SCTccmito",
                                                pt.size = 1,
                                                label = T,
                                                group.by = "seurat_clusters_TCells_SCTccmito_PCA_harmony") +
  theme_jb_nogrid()

umapharmony.SCTccmito.celltypess <- DimPlot(SCC.T,
                                            reduction = "humap_SCTccmito",
                                            pt.size = 1,
                                            label = F,
                                            group.by = "CellType_ss") + 
  theme_jb_nogrid()

umapharmony.SCTccmito.patients <- DimPlot(SCC.T,
                                          reduction = "humap_SCTccmito",
                                          pt.size = 1,
                                          label = F,
                                          group.by = "patient") + 
  theme_jb_nogrid()

umapharmony.SCTccmito.sample <- DimPlot(SCC.T,
                                        reduction = "humap_SCTccmito",
                                        pt.size = 1,
                                        label = F,
                                        group.by = "sample") + 
  theme_jb_nogrid()

umapharmony.SCTccmito.Lok <- DimPlot(SCC.T,
                                     reduction = "humap_SCTccmito",
                                     pt.size = 1,
                                     label = F,
                                     group.by = "Lok") + 
  theme_jb_nogrid()

umapharmony.SCTccmito.lib <- DimPlot(SCC.T,
                                     reduction = "humap_SCTccmito",
                                     pt.size = 1,
                                     label = F,
                                     group.by = "lib") + 
  theme_jb_nogrid()

umapharmony.SCTccmito.Phase <- DimPlot(SCC.T,
                                       reduction = "humap_SCTccmito",
                                       pt.size = 1,
                                       label = F,
                                       group.by = "Phase") + 
  theme_jb_nogrid()

UMAPharmony.SCTccmito.qc <- FeaturePlot(SCC.T, 
                                        features = c("nCount_RNA_log10", "nFeature_RNA_log10", "percent.mito", "n.exp.hkgenes","S.Score","G2M.Score"), 
                                        reduction = "humap_SCTccmito", 
                                        cols = c("grey95", "blue"),
                                        pt.size = 1,
                                        ncol = 2,
                                        order = T,
                                        slot = "data") +
  theme_jb_nogrid()

ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-clusPCAharmony.svg"),umapharmony.SCTccmito.clusPCAharmony)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-clusPCAharmony-noleg.svg"),umapharmony.SCTccmito.clusPCAharmony + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-celltypess.svg"),umapharmony.SCTccmito.celltypess)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-celltypess-noleg.svg"),umapharmony.SCTccmito.celltypess + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-patients.svg"),umapharmony.SCTccmito.patients)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-patients-noleg.svg"),umapharmony.SCTccmito.patients + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-sample.svg"),umapharmony.SCTccmito.sample)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-sample-noleg.svg"),umapharmony.SCTccmito.sample + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-Lok.svg"),umapharmony.SCTccmito.Lok)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-Lok-noleg.svg"),umapharmony.SCTccmito.Lok + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-lib.svg"),umapharmony.SCTccmito.lib)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-lib-noleg.svg"),umapharmony.SCTccmito.lib + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-Phase.svg"),umapharmony.SCTccmito.Phase)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-Phase-noleg.svg"),umapharmony.SCTccmito.Phase + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-QC.svg"),UMAPharmony.SCTccmito.qc, width=10, height=10)

umapharmony.SCTccmito.clusPCAharmony
umapharmony.SCTccmito.celltypess
umapharmony.SCTccmito.patients
umapharmony.SCTccmito.sample
umapharmony.SCTccmito.Lok
umapharmony.SCTccmito.lib
umapharmony.SCTccmito.Phase
UMAPharmony.SCTccmito.qc
```

# Marker Gene Expression

```{r}
UMAPharmony.SCTccmito.marker <- lapply(names(celltypemarkers), function(x) FeaturePlot_markers_comb(SCC.T,celltypemarkers[[x]],reduction = "humap_SCTccmito"))
names(UMAPharmony.SCTccmito.marker) <- names(celltypemarkers)

lapply(names(UMAPharmony.SCTccmito.marker), function(x) ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-marker-",x,".png"),UMAPharmony.SCTccmito.marker[[x]],dpi=300, width=ifelse(length(celltypemarkers[[x]])>1,10,5), height=5*ceiling(length(celltypemarkers[[x]])/2)))

UMAPharmony.SCTccmito.marker
```

# celltype annotation based on marker genes

```{r}
SCC.T$CellType_TCells <- as.character(SCC.T$seurat_clusters_TCells_SCTccmito_PCA_harmony)
SCC.T$CellType_TCells[SCC.T$CellType_TCells=="24"] <- "Mast cells"

SCC.T$CellType_TCells[!SCC.T$CellType_TCells %in% c("Mast cells")] <- "T-Cells"

umapharmony.SCTccmito.celltypetcells <- DimPlot(SCC.T,
                                                reduction = "humap_SCTccmito",
                                                pt.size = 1,
                                                label = T,
                                                group.by = "CellType_TCells") + 
  theme_jb_nogrid()

ggsave(filename = paste0(OUTPUT,"figs/SCCfilt-TCells-SCTccmito-UMAPharmony-CelltypeTCells-noleg.svg"),umapharmony.SCTccmito.celltypetcells + theme(legend.position = "none"), height=5, width = 5)
umapharmony.SCTccmito.celltypetcells
```

#Save

```{r}
#Single Seurat object
save(file = paste0(OUTPUT,"RData/SCCfilt-TCells.RData"),SCC.T)
#SessionInfo
writeLines(capture.output(sessionInfo()), paste0(OUTPUT,"RData/SCCfilt-TCells-sessionInfo.txt"))
```
