---
title: "patient5"
author: "Kai"
date: "6th April 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(Seurat)
library(ggplot2)

#load functions
source(paste0("~/tscr/kai/projects/01_SCC_scRNA/results/functions.R"))
#set output
OUTPUT = "~/tscr/kai/projects/01_SCC_scRNA/results/samples/cohort/patient5/"
```

# Load files

```{r}
data_dir = "~/tscr/Allgemeine Daten/10x Genomics/data/PEI03_A006850046_MCC-SCC/cellranger/SID114998/outs/"

#raw data containing gene expression and Hashtag matrices
patient5.data <- Read10X(paste0(data_dir,"raw_feature_bc_matrix/"))
#Setup Seurat object for gene Expression
patient5 <- CreateSeuratObject(patient5.data, project = "patient5")
#remove raw data input
rm(patient5.data)
```

# Quality control and Filtering

## Assign the expression of mitochondrial and housekeeping genes

houskeeping genes are e.g. ribosomal proteins, expressed in nearly every cell and less sensitive to high dropout.

```{r}
# relative expression of mitochondrial gene as metric for damaged cells
patient5[["percent.mito"]] <- PercentageFeatureSet(patient5, pattern = "^MT-") 

#read in names of housekeeper genes
hkgenes <- read.table("~/tscr/kai/projects/01_SCC_scRNA/db/tirosh_house_keeping.txt", skip = 2)
hkgenes <- as.vector(hkgenes$V1) 
# remove hkgenes that were not found (maybe not in the reference or got lost)
hkgenes.found <- which(toupper(rownames(patient5)) %in% toupper(hkgenes))
# sum of hk genes for each cell
hk_per_cell <- Matrix::colSums(GetAssayData(patient5)[hkgenes.found, ] > 0)
patient5 <- AddMetaData(object = patient5, metadata = hk_per_cell, col.name = "n.exp.hkgenes")
rm(hk_per_cell,hkgenes.found)
```

## Quality Control Plots

Than we have four metrices to identify barcodes containing live and usable cells:
* UMI counts (nCount)
* Gene counts (nFeature)
* Mitochondrial gene expression: % of UMIs in mitochondrial genes relative to all genes
* Housekeeper genes: Number of housekeeping genes (97 in total) that have at least 1 UMI count in a cell (no percentage)

It is not helpful to look at all metrics independently since they correlating a lot with each other. So first, the correlations between the metrices are investigated.

### Filter quality metric dataframe for definitely empty barcodes to get readable histograms

First, we create a helping dataframe and filter out cells that are definitely empty barcodes since they have only very few UMI Counts. 
These are normally around 70-90% of all barcodes. Removing those will give us more informative histograms (Otherwise there will always be a very high peak at around zero). 
To examine distribution over all barcodes, we do a cumulative plot showing the % of cells with less than a certain amount of counts.

```{r cumplot}
#create summary data frame
barcode_metrics <- data.frame(barcodes = names(patient5$nCount_RNA),
                              nCount = patient5$nCount_RNA,
                              nFeature = patient5$nFeature_RNA,
                              percent.mito = patient5$percent.mito,
                              n.exp.hkgenes = patient5$n.exp.hkgenes)

cumplotdata <- data.frame(x=1:1000, y= unlist(lapply(1:1000,function(x) sum(patient5$nCount_RNA < x) / length(patient5$nCount_RNA < x))))

cumplot.log10 <- ggplot() + 
  geom_point(data=cumplotdata,
             aes(x=log10(x),y=y)) +
  ggtitle("% of barcodes that are lower than x Counts") +
  scale_y_continuous(breaks=seq(0,1,0.1)) +
  coord_cartesian(ylim=c(0.5,1)) +
  theme_jb()

ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-cumplot-log10.svg"),cumplot.log10)

cumplot.log10
```

Than we do the actual filtering for around 90% of the cells
```{r}
nCount_lowest <- 5
barcode_metrics <- barcode_metrics[barcode_metrics$nCount > nCount_lowest,]
```

### Examine: UMI counts vs gene counts

UMI and gene counts are highly correlated with each other. It is not possible to have more genes than counts (since every gene need at least one UMI count).

```{r UMIvsGENE}
UMIvsGENE.mito <- PlotUMIvsGENE(barcode_metrics = barcode_metrics,
                                col = barcode_metrics$percent.mito,
                                colname = "percent.mito")
UMIvsGENE.hk <- PlotUMIvsGENE(barcode_metrics = barcode_metrics,
                              col = barcode_metrics$n.exp.hkgenes,
                              colname = "n.exp.hkgenes")

ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsGENE-mito.svg"),UMIvsGENE.mito)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsGENE-hk.svg"),UMIvsGENE.hk)
```

### Examine: mitochondrial gene expression vs housekeeper genes

Usually, barcodes with high mitochondrial gene expression and low housekeeper gene expression are regarded as dying cells or cells undergoing large oxidative stress.

```{r mitvshk}
HKvsMITO.UMI <- PlotHKvsMITO(barcode_metrics = barcode_metrics,
                             col = log10(barcode_metrics$nCount),
                             colname = "nCount_log10")

ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-HKvsMITO-UMI.svg"),HKvsMITO.UMI)
```

## Filtering Counts and Features

```{r}
#Cut-offs
nCount_low <- 500
nCount_high <- NA
nFeature_low <- NA
nFeature_high <- NA

#Subset Seurat object
patient5.filt <- subset(patient5,
                        nCount_RNA > nCount_low
                        #& nCount_RNA < nCount_high
                        #& nFeature_RNA > nFeature_low 
                        #& nFeature_RNA < nFeature_high
)

#assign which barcodes were kept in filtering
barcode_metrics$countfiltering <- NA
barcode_metrics[barcode_metrics$nCount > nCount_low
                #barcode_metrics$nCount < nCount_high & 
                #barcode_metrics$nFeature > nFeature_low &
                #barcode_metrics$nFeature < nFeature_high &
                ,]$countfiltering <- "cell" 
barcode_metrics[!(barcode_metrics$nCount > nCount_low
                  #barcode_metrics$nCount < nCount_high & 
                  #barcode_metrics$nFeature > nFeature_low &
                  #barcode_metrics$nFeature < nFeature_high &
),]$countfiltering <- "background"
```

### Examine: UMI counts vs gene counts after filtering

```{r umivsgene-afterfilt}
#All barcodes, colored by filtered cells with minimum count and feature numbers
UMIvsGENE.filt <- PlotUMIvsGENE(barcode_metrics = barcode_metrics,
                                col = barcode_metrics$countfiltering,
                                colname = "cells filtered", 
                                xintercept = log10(c(nCount_low,nCount_high)), 
                                yintercept = log10(c(nFeature_low,nFeature_high))) 

#only cell barcodes, colored by percent.mito
UMIvsGENE.cells.mito <- PlotUMIvsGENE(barcode_metrics[barcode_metrics$countfiltering == "cell",],
                                      col = barcode_metrics[barcode_metrics$countfiltering == "cell",]$percent.mito,
                                      colname = "percent.mito",
                                      xintercept = log10(c(nCount_low,nCount_high)),
                                      yintercept = log10(c(nFeature_low,nFeature_high)))

# only NOT cells
UMIvsGENE.notcells.mito <-  PlotUMIvsGENE(barcode_metrics[barcode_metrics$countfiltering != "cell",],
                                          col = barcode_metrics[barcode_metrics$countfiltering != "cell",]$percent.mito,
                                          colname = "percent.mito",
                                          xintercept = log10(c(nCount_low,nCount_high)),
                                          yintercept = log10(c(nFeature_low,nFeature_high)))

ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsGENE-filt.svg"),UMIvsGENE.filt)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsGENE-cells-mito.svg"),UMIvsGENE.cells.mito)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsGENE-notcells-mito.svg"),UMIvsGENE.notcells.mito)
```


### Examine: mitochondrial gene expression and housekeeper genes vs counts

```{r mitandhkvsumi-cf}
UMIvsMITO.cf.hk <- PlotUMIvsMITO(barcode_metrics = barcode_metrics[barcode_metrics$countfiltering=="cell",],
                                 col = barcode_metrics[barcode_metrics$countfiltering=="cell",]$n.exp.hkgenes,
                                 colname = "n.exp.hkgenes")
UMIvsHK.cf.mito <- PlotUMIvsHK(barcode_metrics = barcode_metrics[barcode_metrics$countfiltering=="cell",],
                               col = barcode_metrics[barcode_metrics$countfiltering=="cell",]$percent.mito,
                               colname = "percent.mito")

ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsMITO-countfilt-hk.svg"),UMIvsMITO.cf.hk)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsHK-countfilt-mito.svg"),UMIvsHK.cf.mito)
```

### Examine: mitochondrial gene expression vs housekeeper genes

Usually, barcodes with high mitochondrial gene expression and low housekeeper gene expression are regarded as dying cells or cells undergoing large oxidative stress.

```{r mitvshk-cf}
HKvsMITO.cf.UMI <- PlotHKvsMITO(barcode_metrics = barcode_metrics[barcode_metrics$countfiltering=="cell",],
                                col = log10(barcode_metrics[barcode_metrics$countfiltering=="cell",]$nCount),
                                colname = "nCount_log10")

ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-HKvsMITO-countfilt-UMI.svg"),HKvsMITO.cf.UMI)
```

## Filtering all

Complete Filtering including also thresholds for mitochondrial and housekeeper genes.

```{r}
percent.mito_high <- 10
n.exp.hkgenes_low <- 30

#make sure, you also enable the nCount and Feature filter here!
patient5.filt <- subset(patient5,
                        nCount_RNA > nCount_low
                        #& nCount_RNA < nCount_high
                        #& nFeature_RNA > nFeature_low 
                        #& nFeature_RNA < nFeature_high
                        & percent.mito < percent.mito_high 
                        & n.exp.hkgenes > n.exp.hkgenes_low
)

#assign which barcodes were kept in filtering
barcode_metrics$manualfiltering <- NA
barcode_metrics[barcode_metrics$percent.mito < percent.mito_high & 
                  barcode_metrics$n.exp.hkgenes > n.exp.hkgenes_low & 
                  #barcode_metrics$nCount < nCount_high & 
                  #barcode_metrics$nFeature > nFeature_low &
                  #barcode_metrics$nFeature < nFeature_high &
                  barcode_metrics$nCount > nCount_low,]$manualfiltering <- "cell" 
barcode_metrics[!(barcode_metrics$percent.mito < percent.mito_high & 
                    barcode_metrics$n.exp.hkgenes > n.exp.hkgenes_low & 
                    #barcode_metrics$nCount < nCount_high & 
                    #barcode_metrics$nFeature > nFeature_low &
                    #barcode_metrics$nFeature < nFeature_high &
                    barcode_metrics$nCount > nCount_low),]$manualfiltering <- "background"
```

### Examine: UMI counts vs gene counts after filtering

```{r umivsgene-afterfilt}
#All barcodes, colored by filtered cells with minimum count and feature numbers
UMIvsGENE.filt <- PlotUMIvsGENE(barcode_metrics = barcode_metrics,
                                col = barcode_metrics$manualfiltering,
                                colname = "cells filtered", 
                                xintercept = log10(c(nCount_low,nCount_high)), 
                                yintercept = log10(c(nFeature_low,nFeature_high))) 

#only cell barcodes, colored by percent.mito
UMIvsGENE.cells.mito <- PlotUMIvsGENE(barcode_metrics[barcode_metrics$manualfiltering == "cell",],
                                      col = barcode_metrics[barcode_metrics$manualfiltering == "cell",]$percent.mito,
                                      colname = "percent.mito",
                                      xintercept = log10(c(nCount_low,nCount_high)),
                                      yintercept = log10(c(nFeature_low,nFeature_high)))

# only NOT cells
UMIvsGENE.notcells.mito <-  PlotUMIvsGENE(barcode_metrics[barcode_metrics$manualfiltering != "cell",],
                                          col = barcode_metrics[barcode_metrics$manualfiltering != "cell",]$percent.mito,
                                          colname = "percent.mito",
                                          xintercept = log10(c(nCount_low,nCount_high)),
                                          yintercept = log10(c(nFeature_low,nFeature_high)))

ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsGENE-filt.svg"),UMIvsGENE.filt)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsGENE-cells-mito.svg"),UMIvsGENE.cells.mito)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsGENE-notcells-mito.svg"),UMIvsGENE.notcells.mito)
```

### Examine: mitochondrial gene expression and housekeeper genes vs counts after filtering

```{r mitandhkvsumi-afterfilt}
#all barcodes, colored by cells
UMIvsMITO.filt <-  PlotUMIvsMITO(barcode_metrics = barcode_metrics,
                                 col = barcode_metrics$manualfiltering,
                                 colname = "cells filtered",
                                 xintercept = log10(c(nCount_low,nCount_high)),
                                 yintercept = percent.mito_high)
#only cells
UMIvsMITO.cells.hk <-  PlotUMIvsMITO(barcode_metrics = barcode_metrics[barcode_metrics$manualfiltering == "cell",],
                                     col = barcode_metrics[barcode_metrics$manualfiltering == "cell",]$n.exp.hkgenes,
                                     colname = "n.exp.hkgenes",
                                     xintercept = log10(c(nCount_low,nCount_high)),
                                     yintercept = percent.mito_high) 
#only not cells
UMIvsMITO.notcells.hk <-  PlotUMIvsMITO(barcode_metrics = barcode_metrics[barcode_metrics$manualfiltering != "cell",],
                                        col = barcode_metrics[barcode_metrics$manualfiltering != "cell",]$n.exp.hkgenes,
                                        colname = "n.exp.hkgenes",
                                        xintercept = log10(c(nCount_low,nCount_high)),
                                        yintercept = percent.mito_high) 

#all barcodes, colored by cells
UMIvsHK.filt <-  PlotUMIvsHK(barcode_metrics = barcode_metrics,
                             col = barcode_metrics$manualfiltering,
                             colname = "cell",
                             xintercept = log10(c(nCount_low,nCount_high)),
                             yintercept = n.exp.hkgenes_low)
#only cells
UMIvsHK.cells.mito <- PlotUMIvsHK(barcode_metrics = barcode_metrics[barcode_metrics$manualfiltering == "cell",],
                                  col = barcode_metrics[barcode_metrics$manualfiltering == "cell",]$percent.mito,
                                  colname = "percent.mito",
                                  xintercept = log10(c(nCount_low,nCount_high)),
                                  yintercept = n.exp.hkgenes_low) 

#only not cells
UMIvsHK.notcells.mito <- PlotUMIvsHK(barcode_metrics = barcode_metrics[barcode_metrics$manualfiltering != "cell",],
                                     col = barcode_metrics[barcode_metrics$manualfiltering != "cell",]$percent.mito,
                                     colname = "percent.mito",
                                     xintercept = log10(c(nCount_low,nCount_high)),
                                     yintercept = n.exp.hkgenes_low) 

ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsMITO-filt.svg"),UMIvsMITO.filt)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsMITO-cells-hk.svg"),UMIvsMITO.cells.hk)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsMITO-notcells-hk.svg"),UMIvsMITO.notcells.hk)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsHK-filt.svg"),UMIvsHK.filt)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsHK-cells-mito.svg"),UMIvsHK.cells.mito)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-UMIvsHK-notcells-mito.svg"),UMIvsHK.notcells.mito)
```

### Examine: percent.mito vs n.exp.hkgenes after filtering

```{r HKvsMITO-afterfilt}
#all barcodes, colored by filtering
HKvsMITO.filt <- PlotHKvsMITO(barcode_metrics = barcode_metrics,
                              col = barcode_metrics$manualfiltering,
                              colname = "cell",
                              xintercept = n.exp.hkgenes_low,
                              yintercept = percent.mito_high)

#only cells, colored by nCounts
HKvsMITO.cells.UMI <- PlotHKvsMITO(barcode_metrics = barcode_metrics[barcode_metrics$manualfiltering == "cell",],
                                   col = log10(barcode_metrics[barcode_metrics$manualfiltering == "cell",]$nCount),
                                   colname = "nCount_log10",
                                   xintercept = n.exp.hkgenes_low, 
                                   yintercept = percent.mito_high)

#only NOT cells, colored by nCounts
HKvsMITO.notcells.UMI <- PlotHKvsMITO(barcode_metrics = barcode_metrics[barcode_metrics$manualfiltering != "cell",],
                                      col = log10(barcode_metrics[barcode_metrics$manualfiltering != "cell",]$nCount),
                                      colname = "nCount_log10",
                                      xintercept = n.exp.hkgenes_low, 
                                      yintercept = percent.mito_high)

ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-HKvsMITO-filt.svg"),HKvsMITO.filt)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-HKvsMITO-cells-UMI.svg"),HKvsMITO.cells.UMI)
ggsave(filename = paste0(OUTPUT,"figs/patient5-filtering-HKvsMITO-notcells-UMI.svg"),HKvsMITO.notcells.UMI)
```

# Normalization

## Normalization of RNA assay

```{r}
#log-normalization for slot "data"
patient5.filt <- NormalizeData(patient5.filt, 
                               normalization.method = "LogNormalize", #default
                               scale.factor = 10000, #default, scale factor 10000 is recommendation from seurat tutorial
                               assay = "RNA",
                               margin = 1 # default; normalizes across features
) 

#Find Variable Features
patient5.filt <- FindVariableFeatures(patient5.filt, 
                                      assay = "RNA",
                                      selection.method = "vst", #default
                                      nfeatures = 2000 #default; only 2000 , 3000 for SCT promises since to "restore" some more VariableFeatures
)

#scaling for slot "scale.data" by all genes, not only most variable ones
allgenes <- rownames(patient5.filt)
patient5.filt <- ScaleData(patient5.filt, 
                           features = allgenes, 
                           do.scale = T,
                           do.center = T,
                           scale.max = 10, 
                           assay = "RNA")
```

## Determine cell cycle

```{r}
patient5.filt <- CellCycleScoring(patient5.filt,
                                  s.features = cc.genes.updated.2019$s.genes,
                                  g2m.features = cc.genes.updated.2019$g2m.genes, 
                                  assay="RNA")
```

## Normalization with SCTransform

```{r}
patient5.filt <- SCTransform(patient5.filt, 
                             ncells = ncol(patient5.filt),
                             assay="RNA", #default
                             new.assay.name = "SCT", #default
                             do.correct.umi = T, #default change counts slot to corrected counts in new assay
                             variable.features.n = 3000, #default set variable features
                             vars.to.regress = NULL, #default optional variables to regress out
                             do.scale = F, #default scale pearson residuals in scale.data slot to have unit variance
                             do.center = T, #default center pearson residuals in scale.data to have mean zero, needed for PCA
                             return.only.var.genes = T, #default scale.data.matrices output assay only contains variable genes
                             seed.use = 1448145)
```

# Dimension reduction

## PCA

```{r}
patient5.filt <- RunPCA(patient5.filt, 
                        npcs = 50,  #number of PCs to use
                        assay = "SCT",
                        rev.pca = F, # default, Run cell x gene matrix
                        weight.by.var = T, #default, Weight cell embedding by variance of each PC!)
                        approx = T, #if TRUE, uses irlba instead of prcomp, which is a single value decomposition to approximate PCs, saving computation time (by not calculating all PCs; important for big datasets) and being similar/same to PCs
                        features = VariableFeatures(patient5.filt,assay = "SCT"), #default
                        reduction.name = "pca_SCT",
                        verbose = F)
```

```{r}
#calculate the % variance explained
PCAvar.SCT <- patient5.filt@reductions$pca_SCT@stdev^2 / sum(matrixStats::rowVars(as.matrix(patient5.filt@assays$SCT@scale.data)[VariableFeatures(patient5.filt,assay = "SCT"),]))

#PCA plot
PCAplot.SCT <- DimPlot(patient5.filt, reduction = "pca_SCT", dims = c(1,2), group.by = "orig.ident") + 
  labs(x=paste0("principal component 1 (",signif(PCAvar.SCT[1]*100,digits = 2),"% variance)"),
       y=paste0("principal component 2 (",signif(PCAvar.SCT[2]*100,digits = 2),"% variance)")) +
  theme_jb()

#elbow plot
PCAelbowplot.SCT <- ggplot(data=data.frame(var=PCAvar.SCT,
                                           PC=1:length(PCAvar.SCT))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=var)) +
  labs(x="principal component", y="% variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

# cumulative Variance Plot
PCAsumplot.SCT <- ggplot(data=data.frame(cumvar=cumsum(PCAvar.SCT),PC=1:length(PCAvar.SCT))[1:50,]) +
  geom_point(mapping = aes(x=PC,y=cumvar)) +
  labs(x="principal component", y="% cumulative variance") +
  scale_x_continuous(breaks=seq(0,50,5)) +
  theme_jb()

ggsave(filename = paste0(OUTPUT,"figs/patient5-all-SCT-PCA.svg"),PCAplot.SCT, width = 5, height = 5)
# without legends
ggsave(filename = paste0(OUTPUT,"figs/patient5-all-SCT-PCA-noleg.svg"),PCAplot.SCT + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/patient5-all-SCT-PCA-elbow.svg"),PCAelbowplot.SCT, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/patient5-all-SCT-PCA-elbow-noleg.svg"),PCAelbowplot.SCT + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/patient5-all-SCT-PCA-variancesum.svg"),PCAsumplot.SCT, width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/patient5-all-SCT-PCA-variancesum-noleg.svg"),PCAsumplot.SCT + theme(legend.position = "none") + labs(title = NULL),width = 5, height = 5)

PCAplot.SCT
PCAelbowplot.SCT
PCAsumplot.SCT
```

## Graph-based clustering

```{r FindClusters}
numberofPCs = 15

#based on PCA
patient5.filt <- FindNeighbors(patient5.filt, 
                               dims = 1:numberofPCs, 
                               reduction = "pca_SCT",
                               assay = "SCT")
patient5.filt <- FindClusters(patient5.filt, 
                              resolution = 1,
                              random.seed=100)

patient5.filt$seurat_clusters_all_SCT_PCA <- patient5.filt$seurat_clusters
patient5.filt$seurat_clusters <- NULL 
```

## UMAP

```{r UMAP, fig.width=10, fig.height=8}
patient5.filt <- RunUMAP(patient5.filt,
                         dims = 1:numberofPCs,
                         assay = "SCT",
                         umap.method = "uwot", # Seurat default
                         graph=NULL, #default
                         reduction="pca_SCT",
                         reduction.name = "umap_SCT"
)
```

Explore UMAP with different annotations

```{r}
UMAP.SCT.clusPCA <- DimPlot(patient5.filt, 
                            reduction = "umap_SCT", 
                            pt.size = 1, 
                            label = T, 
                            group.by = "seurat_clusters_all_SCT_PCA") +
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()

UMAP.SCT.QC <- FeaturePlot(patient5.filt, 
                           features = c("nCount_RNA_log10", "nFeature_RNA_log10", "percent.mito", "n.exp.hkgenes","S.Score","G2M.Score"), 
                           reduction = "umap_SCT", 
                           cols = c("grey95", "blue"),
                           pt.size = 1,
                           order = T,
                           slot = "data")

ggsave(filename = paste0(OUTPUT,"figs/patient5-all-SCT-UMAP-clusPCA-noleg.svg"),UMAP.SCT.clusPCA + theme(legend.position = "none") + labs(title = NULL), width = 5, height = 5)
ggsave(filename = paste0(OUTPUT,"figs/patient5-all-SCT-UMAP-QC.svg"),UMAP.SCT.QC, width=10, height=15)

UMAP.SCT.clusPCA
UMAP.SCT.QC
```

# Cell Type identification

Manual annotation based on marker gene expression from various literature references.

## Marker Gene Expression

```{r}
UMAP.SCT.marker <- lapply(names(celltypemarkers), function(x) FeaturePlot_markers_comb(patient5.filt,celltypemarkers[[x]],reduction = "umap_SCT"))
names(UMAP.SCT.marker) <- names(celltypemarkers)

lapply(names(UMAP.SCT.marker), function(x) ggsave(filename = paste0(OUTPUT,"figs/patient5-all-SCT-UMAP-marker-",x,".svg"),UMAP.SCT.marker[[x]], width=ifelse(length(celltypemarkers[[x]])>1,10,5), height=5*ceiling(length(celltypemarkers[[x]])/2)))

UMAP.SCT.marker
```

## Celltype Annotation

```{r umap-celltype}
Idents(patient5.filt) <- patient5.filt$seurat_clusters_all_SCT_PCA
patient5.filt <- RenameIdents(patient5.filt,
                              "0" = "Immune cells",
                              "1" = "stroma cells"
)
patient5.filt$CellType <- Idents(patient5.filt)
```

## UMAP

```{r}
UMAP.SCT.celltype <- DimPlot(patient5.filt,
                             reduction = "umap_SCT", 
                             pt.size = 1, 
                             label = T, 
                             group.by = "CellType") + 
  labs(x="UMAP_1",y="UMAP_2") +
  theme_jb_nogrid()
ggsave(filename = paste0(OUTPUT,"figs/patient2-all-SCT-UMAP-celltype.svg"),UMAP.SCT.celltype, width=5, height=5)
ggsave(filename = paste0(OUTPUT,"figs/patient2-all-SCT-UMAP-celltype-noleg.svg"),UMAP.SCT.celltype + theme(legend.position = "none") + labs(title = NULL), width=5, height=5)
UMAP.SCT.celltype
```

# Save 

```{r}
#single Seurat object
save(file = paste0(OUTPUT,"RData/patient5.RData"),patient5.filt)
#sessionInfo
writeLines(capture.output(sessionInfo()), paste0(OUTPUT,"RData/patient5-preprocessing-sessionInfo.txt"))
```
